### 2020-04-01 research-cv/research-cv.tex
2 page

%Research CV template for academics and researchers
%@TheDoctorRAB
\documentclass[letter,oneside,english,10pt,headinclude,headsepline]{amsart}
\usepackage[margin=1in]{geometry}
\usepackage{times}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{bold-extra}
\usepackage{bibentry}
\usepackage{filecontents}
\usepackage[yyyymmdd]{datetime} %set preferred date format
\renewcommand{\dateseparator}{.} %set preferred date separator

\makeatletter %makes section in block caps with large font \section*{}
\renewcommand
\section{\@startsection{section}{1}%
  \z@{1.5\linespacing}{.25\linespacing}%
  {\bfseries\scshape\large\raggedright}}

\makeatletter %makes section in block caps with normal font \subsection*{}
\renewcommand
\subsection{\@startsection{subsection}{1}%
  \z@{.75\linespacing}{.25\linespacing}%
  {\bfseries\scshape\raggedright}}

\newcommand*\sq{\mathbin{\vcenter{\hbox{\rule{.3ex}{.3ex}}}}} %makes a small square as a separator $\sq$

\usepackage{fancyhdr} 
\pagestyle{fancy}
\fancyhf{} %page number on bottom right
\lhead{\scriptsize Title Name} %title and name
\chead{\scriptsize \textit{Vita}} %maybe something more specific here if desired
\rhead{\scriptsize \today} %date
\rfoot{\thepage} %page number

\title{Name}

%peer reviewed journal papers
\begin{filecontents}{peer-reviewed.bib} 
    @article{
        pr1,
        author = {},
        journal = {},
        pages = {},
        title = {{}},
        volume = {},
        year = {}
    }
\end{filecontents}

%books and book contributions
\begin{filecontents}{book-contribution.bib}
    @book{
        bk1,
        author = {},
        title = {{}},
        year = {},
        publisher = {{}}
        isbn = {}
    }
\end{filecontents}

%conference proceedings
\begin{filecontents}{conference-proceedings.bib}
    @conference{
        cf1,
        author = {},
        title = {{}},
        year = {},
        organization = {},
        address = {}
    }
\end{filecontents}

%reports, misc, non-refeered
\begin{filecontents}{reports.bib}
    @misc{
        rp1,
        author = {},
        title = {{}},
        institution = {},
        year = {}
    }
\end{filecontents}

%conference presentations - not the same as invited talks 
\begin{filecontents}{conference-presentations.bib}
    @conference{
        cf1,
        author = {},
        title = {{}},
        year = {},
        organization = {},
        address = {}
    }
\end{filecontents}

\begin{document}

%.bib files to create
\nobibliography{peer-reviewed,book-contribution,conference-proceedings,reports,conference-presentations}

%.bst file
\bibliographystyle{} %custom bst file

\thispagestyle{fancy}

\begin{center}{
        \textsc{ %block caps font
            \textbf{
                {\large Name}\\
                Title \\
                Institution $\sq$ Department
                Address
                email $\sq$ twitter $\sq$ phone 
            }
        }
    }
\end{center}

\noindent\makebox[\linewidth]{\rule{\textwidth}{0.5pt}} %horizontal line spanning margins

\section*{Education}
\noindent\textbf{University}\\
\textbf{Doctor of Philosophy $\sq$ Major} \hfill{\textit{Date}}
\begin{itemize}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
    \item[]\textit{Dissertation: }
    \item[]Examination fields: 
    \item[]Committee: 
\end{itemize}
\vspace*{.75\baselineskip}

\noindent\textbf{University}\\
\textbf{Master of Science $\sq$ Major} \hfill{\textit{Date}}
\begin{itemize}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
    \item[]\textit{Thesis: }
    \item[]Advisor: 
\end{itemize}
\vspace*{.75\baselineskip}

\noindent\textbf{University}\\
\textbf{Bachelor of Science $\sq$ Major and distinction(s)} \hfill{\textit{}}
\begin{itemize}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
    \item[]\textit{Capstone: }
    \item[] Advisor: 
\end{itemize}

\section*{Research Experience} %academic positions
\noindent\textbf{University/Institution $\sq$ Location $\sq$ Department}\\
\textbf{Title} \hfill{\textit{From - To}}\\
Brief summary 
\vspace*{.75\baselineskip}

\noindent\textbf{University/Institution $\sq$ Location $\sq$ Department}\\
\textbf{Title} \hfill{\textit{From - To}}\\
Brief summary

\section*{Related Research Experience} %like internships
\noindent\textbf{University/Institution $\sq$ Location $\sq$ Department}\\
\textbf{Title} \hfill{\textit{From - To}}\\
Brief summary 
\vspace*{.75\baselineskip}

\noindent\textbf{University/Institution $\sq$ Location $\sq$ Department}\\
\textbf{Title} \hfill{\textit{From - To}}\\
Brief summary

\section*{Related Professional Experience} %anything else
\noindent\textbf{University/Institution $\sq$ Location $\sq$ Department}\\
\textbf{Title} \hfill{\textit{From - To}}\\
Brief summary 
\vspace*{.75\baselineskip}

\noindent\textbf{University/Institution $\sq$ Location $\sq$ Department}\\
\textbf{Title} \hfill{\textit{From - To}}\\
Brief summary

\section*{Awarded Projects} %chronologically
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=.15ex,partopsep=1ex,parsep=1ex]
    \item
    \item
    \item
\end{enumerate}

\section*{Thesis Committees}
\subsection*{Major Professor} %or adviser, whatever the official title is
\noindent\textit{Current}
\begin{itemize}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
    \item[]Student Name, Degree - Topic
\end{itemize}
\vspace*{.75\baselineskip}

\noindent\textit{Graduates}
\begin{itemize}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
    \item[]Student Name, Degree - Thesis Title 
\end{itemize}

\subsection*{Member}
\noindent\textit{Graduates}
\begin{itemize}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
    \item[]Student Name, Degree - Thesis Title 
\end{itemize}

\section*{Refereed Journal Publications}
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=.15ex,partopsep=1ex,parsep=1ex]
    \item\bibentry{pr1}.
\end{enumerate}

\section*{Book Contribution}
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=.15ex,partopsep=1ex,parsep=1ex]
    \item\bibentry{bk1}.
\end{enumerate}

\section*{Refereed Conference Proceedings}
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=.15ex,partopsep=1ex,parsep=1ex]
    \item\bibentry{cf1}.
\end{enumerate}

\section*{Non-Refereed Technical Reports}
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=.15ex,partopsep=1ex,parsep=1ex]
    \item\bibentry{rp1}.
\end{enumerate}

\section*{Conference Presentations}
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=.15ex,partopsep=1ex,parsep=1ex]
    \item\bibentry{cf1}.
\end{enumerate}

\section*{Teaching Experience}
\noindent\textbf{University/Institution $\sq$ Location $\sq$ Department}\\
\textbf{Title}\\ 
\textit{Course Number: Title \hfill{Semester}}\\
Summary
\vspace*{.75\baselineskip}

\noindent\textbf{University/Institution $\sq$ Location $\sq$ Department}\\
\textbf{Title}\\ 
\textit{Course Number: Title \hfill{Semester}}\\
Summary

\section*{Related Teaching Experience} %for TA, etc.
\noindent\textbf{University/Institution $\sq$ Location $\sq$ Department}\\
\textbf{Title}\\ 
\textit{Course Number: Title \hfill{Semester}}\\
Summary
\vspace*{.75\baselineskip}

\noindent\textbf{University/Institution $\sq$ Location $\sq$ Department}\\
\textbf{Title}\\ 
\textit{Course Number: Title \hfill{Semester}}\\
Summary

\section*{Professional Service} 
\subsection*{University/Institution} %department/college related activities
\begin{itemize}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
    \item[]
\end{itemize}

\subsection*{Academic Search Committees}
\begin{itemize}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
    \item[]\textit{Year}
    \item[]Search committee %don't say it was a failed search 
            \vspace{.05in}
    \item[]\textit{Year}
    \item[]Search committee %don't say it was a failed search 
\end{itemize}

\subsection*{Peer Reviewer}
\begin{itemize}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
    \item[]\textit{Regular periods} %for journals/conferences on a regular basis
    \item[] 
            \vspace{.05in}
    \item[]\textit{Year} %specific one-time reviewing
    \item[] 
            \vspace{.05in}
    \item[]\textit{Year} %specific one-time reviewing
    \item[] 
\end{itemize}

\subsection*{Conference Committees}
\begin{itemize}[leftmargin=*,topsep=0pt,itemsep=.05ex,partopsep=1ex,parsep=1ex]
    \item[]
\end{itemize}

\subsection*{Technical Session Organizer}
\begin{itemize}[leftmargin=*,topsep=0pt,itemsep=.05ex,partopsep=1ex,parsep=1ex]
    \item[]
\end{itemize}

\subsection*{Technical Session Chair}
\begin{itemize}[leftmargin=*,topsep=0pt,itemsep=.05ex,partopsep=1ex,parsep=1ex]
    \item[]
\end{itemize}

\subsection*{Symposia Organizer}
\begin{itemize}[leftmargin=*,topsep=0pt,itemsep=.05ex,partopsep=1ex,parsep=1ex]
    \item[]
\end{itemize}

\subsection*{Invited Talks}
\begin{itemize}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
    \item[]\textit{University/Institution $\sq$ Department} %speaking at graduate orientation, etc.
    \item[]Title $\sq$ Purpose \hfill{\textit{Dates}}
        \vspace{.05in}
    \item[]\textit{Seminars} %typical department seminar talks
    \item[]University/Institution Department \hfill{\textit{Date}}
        \vspace{.05in}
    \item[]\textit{Lightning Talks} %throw everything in there 
    \item[]Talk \hfill{\textit{Date}}
        \vspace{.05in}
    \item[]\textit{Outreach}
    \item[]Event \hfill{\textit{Date}}
        \vspace{.05in}
    \item[]\textit{Lectures} %invited to lecture in a course
    \item[]University/Institution - Course or topic \hfill{\textit{Date}}
\end{itemize}

\subsection*{Community Service}
\begin{itemize}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
    \item[]\textit{Event} \hfill{\textit{Date}}
    \item[]Summary 
        \vspace{.05in}
    \item[]\textit{Event} \hfill{\textit{Date}}
    \item[]Summary 
\end{itemize}

\subsection*{Outreach and Related Professional Activities} %career fairs, etc.
\begin{itemize}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
    \item[]Event \hfill{\textit{Date}}
\end{itemize}

\subsection*{Professional Training}
\noindent\textbf{University/Institution}\\
\textbf{Role}\\
\textit{Title of Event} \hfill{\textit{Date}}\\
Summary
\vspace*{.75\baselineskip}

\noindent\textbf{University/Institution}\\
\textbf{Role}\\
\textit{Title of Event} \hfill{\textit{Date}}\\
Summary

\subsection*{Awards and Fellowships}
\begin{itemize}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
    \item[]\textbf{University/Institution}
    \item[]Award or fellowship \hfill{\textit{Date}}
        \vspace{.05in}
    \item[]\textbf{University/Institution}
    \item[]Award or fellowship \hfill{\textit{Date}}

\subsection*{Professional Societies and Positions}
\begin{itemize}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
    \item[]\textbf{Society}
    \item[]Position \hfill{\textit{From - To}}
        \vspace{.05in}
    \item[]\textbf{Society}
    \item[]Position \hfill{\textit{From - To}}
\end{itemize}

\end{document}