### 2020-04-22 journal/journal-paper-standard-coverletter.tex
journal paper

%TheDoctorRAB
%Used for first draft or revised manuscript
\documentclass[11pt,a4paper]{article}
\usepackage[lmargin=1in,rmargin=1in,tmargin=1in,bmargin=1in]{geometry}
\usepackage{setspace}
\usepackage{ulem} %strikethrough - do not \sout{\cite{}}
\usepackage{xcolor} %change font color
\usepackage{filecontents}
\usepackage[yyyymmdd]{datetime}
\renewcommand{\dateseparator}{.}
\usepackage{times}
\usepackage[stable,hang,flushmargin]{footmisc} %footnotes in section titles and no indent
\usepackage{enumitem}
\usepackage{boldline}

\usepackage{xr}
\externaldocument{} %aux file needed for cross references

\newcommand*\sq{\mathbin{\vcenter{\hbox{\rule{.3ex}{.3ex}}}}} %makes a small square as a separator $\sq$

\newcommand{\edit}[1]{\textcolor{blue}{#1}} %shortcut for changing font color on revised text
\newcommand{\fn}[1]{\footnote{#1}} %shortcut for footnote tag

%%% move page number to bottom right
\usepackage{fancyhdr}
\fancyhf{}
\renewcommand{\headrulewidth}{0pt}
\rfoot{\thepage}
\pagestyle{fancy}
%%%

\begin{document}

\thispagestyle{empty} %start with page number 1 on second page

\begin{flushright}
    \today \\
    Lead author contact information
\end{flushright}

\noindent Dr. Editor Name \\
Editor-in-Chief \\
\textit{Journal} \\ \\

\noindent Dear Dr. Editor Name: \\

\noindent We would like to submit the (revised, original) manuscript to \textit{Journal} entitled, `Fancy Title'. \\

\noindent (Each of the three reviewer comments were studied in detail, and a careful response was prepared for each comment. In the marked manuscript, omissions are recorded \sout{as follows}. For revisions, \edit{colored font was used}, and a footnote was provided to indicate the comment from the reviewer.\fn{R1 - Comment from Reviewer 1.})\\

\noindent Thank you (again) for this opportunity. Please refer to my contact information above.

\vspace*{\fill}
\begin{flushright}
    Sincerely,\\
    Lead author
\end{flushright}
\noindent \small{encl.: Marked Manuscript, Clean Manuscript, Response to Reviewer 1, Response to Reviewer 2, Response to Reviewer 3} %only major documents

\end{document}