### 2020-08-18 proposals/white-paper/white-paper.tex
white paper update

%@TheDoctorRAB
%Standard white paper/preproposal format when there is a page limit
%neup.bst designed to use numbered citations in order of appearance, short author list
\documentclass[11pt,letterpaper]{article}
\usepackage[lmargin=1in,rmargin=1in,tmargin=1in,bmargin=1in]{geometry}
\usepackage[pagewise]{lineno} %line numbering
\usepackage{setspace}
\usepackage{ulem} %strikethrough
\usepackage{xcolor} %change font color
\usepackage{graphicx}
\usepackage{filecontents}
\usepackage{tablefootnote}
\usepackage{subfig}
\usepackage[yyyymmdd]{datetime}
\renewcommand{\dateseparator}{.}
\graphicspath{{../img/}}
\setcounter{secnumdepth}{5} %set subsection to nth level
\usepackage{caption}
\captionsetup[table]{skip=0pt} %sets a space after table caption
\usepackage{times}
\usepackage{enumitem}
%\usepackage[labelfont=bf]{caption}

\usepackage[numbers]{natbib} %use 'numbers' for numbered citations; 'round' for () instead [] for inline citations
\setlength{\bibsep}{0pt} %sets space between references
\renewcommand{\bibsection}{} %suppresses large 'references' heading
\renewcommand\bibpreamble{\vspace{-0.2\baselineskip}} %sets spacing after heading if not using default references heading

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{} %move page number to bottom right
\renewcommand{\headrulewidth}{0.5pt} %turn off line in header
\lhead{\scriptsize Name}
\chead{\scriptsize Title, subject, agency, etc.}
\rhead{\scriptsize \today}
\rfoot{\thepage}

\begin{filecontents}{references.bib}
\end{filecontents}

\begin{document}

\begin{center}\textbf{Title}\end{center}\\

\noindent\textbf{Summary of the Proposed Project.} 
\\

\noindent\textbf{Overall Objective.}
\\

\noindent\textbf{Importance and Relevance to Objectives.}
\\

\noindent\textbf{Logical Path, Work Scope, Description of Tasks.} The Logical Path and the proposed Work Scope is described below in terms of the defined Tasks for the project.
\\

\noindent\textbf{Task I. Title.}
\\

\noindent\textbf{Task II. Title.}
\\

\noindent\textbf{Task III. Title.}
\\

\noindent\textbf{Task N. Title.}

\begin{table}[h!]
    \centering
    \caption{Timeframe for Execution of Proposed Project.}
    \begin{tabular}{|l|l|l|l|l|l|}
        \hline
        \textbf{Tasks} & \textbf{M} & \textbf{M} & \textbf{M} & \textbf{M} & \textbf{M} \\
        \hline
        Task I & & & & & \\
        \hline
        Task II & & & & & \\
        \hline
        Task III & & & & & \\
        \hline
    \end{tabular}
    \label{tab-timeframe}
\end{table}

\noindent\textbf{References}
\bibliographystyle{neup}
\setlength{\bibhang}{0pt}
\bibliography{references}

\end{document}