### 2020-12-02 research-cv/abbreviated-research-two-page-cv.tex

%@TheDoctorRAB
\documentclass[letter,oneside,english,10pt,headinclude,headsepline]{amsart}
\usepackage[margin=1in]{geometry}
\usepackage{times}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{bold-extra}
\usepackage{bibentry}
\usepackage{filecontents}
\usepackage[yyyymmdd]{datetime}
\renewcommand{\dateseparator}{.}

\makeatletter %makes section in block caps with large font \section*{}
\renewcommand
\section{\@startsection{section}{1}%
  \z@{1.5\linespacing}{.25\linespacing}%
  {\bfseries\scshape\large\raggedright}}

\makeatletter %makes section in block caps with normal font \subsection*{}
\renewcommand
\subsection{\@startsection{subsection}{1}%
  \z@{.75\linespacing}{.25\linespacing}%
  {\bfseries\scshape\raggedright}}

\newcommand*\sq{\mathbin{\vcenter{\hbox{\rule{.3ex}{.3ex}}}}} %makes a small square as a separator $\sq$

\usepackage{fancyhdr} 
\pagestyle{fancy}
\fancyhf{} %page number on bottom right
\lhead{\scriptsize Title Name}
\chead{\scriptsize \textit{Abbrv. Vita}}
\rhead{\scriptsize \today} %date
\rfoot{\thepage} %page number

\title{Name}

%\natexlab not working - remove from bbl file
\begin{filecontents}{peer-reviewed.bib} 
    @article{
        pr1,
        author = {},
        journal = {},
        pages = {},
        title = {{}},
        volume = {},
        year = {}
    }
\end{filecontents}

\begin{document}

\nobibliography{peer-reviewed} %bib file
\bibliographystyle{standard} %custom bst file

\thispagestyle{fancy}

{\centering 
        \textsc{ %block caps font
            \textbf{
                Name $\sq$ Title \\
                University/Institution $\sq$ Department \\
                Address 1 $\sq$ Address 2\\
                email $\sq$ @Twitter $\sq$ office phone 
            }
        }
\par}

\noindent\makebox[\linewidth]{\rule{\textwidth}{0.5pt}} %horizontal line spanning margins

\subsection*{Education and Training}
\noindent University $\sq$ Doctor of Philosophy $\sq$ Major \hfill{Date}\\
University $\sq$ Master of Science $\sq$ Major \hfill{Date}\\
University $\sq$ Bachelor of Science $\sq$ Major \hfill{Date}

\subsection*{Research and Professional Experience}
\noindent University $\sq$ Location $\sq$ Department
\begin{itemize}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
    \item[]Title/Position \hfill{From - To}
\end{itemize}
\vspace*{.5\baselineskip}

\noindent University $\sq$ Location $\sq$ Department
\begin{itemize}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
    \item[]Title/Position \hfill{From - To}
\end{itemize}

\subsection*{Awarded Projects} %chronologically
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=.05ex,partopsep=1ex,parsep=1ex]
    \item
    \item
    \item
\end{enumerate}

\subsection*{Relevant Publications\hfill{*students}} 
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=.05ex,partopsep=1ex,parsep=1ex]
    \item\bibentry{pr1}.
\end{enumerate}

\subsection*{Relevant Courses Taught}
\begin{itemize}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
    \item[]University $\sq$ Location $\sq$ Department 
    \item[] Course Number: Title 
        \vspace{.05in}
    \item[]University $\sq$ Location $\sq$ Department 
    \item[] Course Number: Title 
\end{itemize}

\subsection*{Synergistic Activities}
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
    \item
    \item
    \item
    \item
    \item
\end{enumerate}

\end{document}