### 2021-01-08 proposal-documents/doe/narrative/doe-proposal-narrative.tex

\documentclass[11pt,letterpaper]{article}
\usepackage[lmargin=1in,rmargin=1in,tmargin=1in,bmargin=1in]{geometry}
\usepackage[pagewise]{lineno} %line numbering
\usepackage{setspace}
\usepackage{ulem} %strikethrough
\usepackage{xcolor,colortbl} %change font color
\usepackage{graphicx}
\usepackage{filecontents}
\usepackage{tablefootnote}
\usepackage{subfig}
\usepackage[yyyymmdd]{datetime}
\renewcommand{\dateseparator}{.}
\graphicspath{{../img/}}
\setcounter{secnumdepth}{5} %set subsection to nth level
\usepackage{times}
\usepackage{enumitem}
\usepackage{float}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage{tabulary}
\usepackage{gensymb}
\usepackage{longtable}
%\usepackage[labelfont=bf]{caption}

\usepackage[singlelinecheck=false]{caption}
\captionsetup[table]{skip=0pt} %sets a space after table caption
\captionsetup[figure]{skip=0pt,labelformat={default},labelsep=period,name={Fig.}} %sets space above caption, 'figure' format

\usepackage{wrapfig}
\setlength{\intextsep}{0.20mm}
\setlength{\columnsep}{0.20mm}

\usepackage[numbers,sort&compress]{natbib} 
%'numbers' - numbered citations
%'round' - (1) instead [1]
%'sort&compress - [1-3] instead [1,2,3]
\setlength{\bibsep}{0pt} %sets space between references
\renewcommand{\bibsection}{} %suppresses large 'references' heading
\renewcommand\bibpreamble{\vspace{-0.2\baselineskip}} %sets spacing after heading if not using default references heading

\newcommand{\edit}[1]{\textcolor{blue}{#1}} %shortcut for changing font color on revised text
\newcommand{\sk}[1]{\sout{#1}} %shortcut for strikethrough
\newcommand{\fn}[1]{\footnote{#1}} %shortcut for footnote tag
\newcommand*\sq{\mathbin{\vcenter{\hbox{\rule{.3ex}{.3ex}}}}} %makes a small square as a separator $\sq$
\newcommand{\x}{\cellcolor{lightgray}} %use to shade in table cell

\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}p{#1}} %uses \raggedright with p{} in table column

\makeatletter
\renewcommand\paragraph{%
    \@startsection{paragraph}{4}{\z@ }{0.35\baselineskip}{-1em}
    {\normalfont \normalsize \bfseries}}%

\makeatletter
\renewcommand\subparagraph{%
    \@startsection{subparagraph}{5}{\z@ }{0.20\baselineskip}{-1em}
    {\normalfont \normalsize \itshape }}%

\makeatletter
\renewcommand\subsection{%
    \@startsection{subsection}{2}{\z@ }{0.50\baselineskip}{0.10\baselineskip}
    { \normalfont \bfseries}}%

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{} %move page number to bottom right
\renewcommand{\headrulewidth}{0pt} %set line thickness in header
%\lhead{\scriptsize Name}
%\chead{\scriptsize Title, subject, agency, etc.}
%\rhead{\scriptsize \today}
\rfoot{\thepage}

%cite in order of appearance if using neup.bst
\begin{filecontents}{references.bib} 
\end{filecontents}

\begin{document}

{\centering 
    \textbf{Title\\
    Call/Technical Scope\\
    }
    Name (PI) - Affiliation\\
    Name (co-PI) - Affiliation
\par
}

\subsection*{Summary of the Proposed Project}

%\begin{wrapfigure}{l}{0.42\textwidth}
%    \includegraphics[width=0.40\textwidth]{flowsheet.jpg}
%    \captionsetup{justification=centering,belowskip=-0.7mm}
%    \caption{Pyroprocessing flowsheet.}
%    \label{fig-flowsheet}
%\end{wrapfigure}

\subsection*{Motivation} 
\paragraph*{Context.}

\paragraph*{Something else.}

\subsection*{Overall Objective}

\subsection*{Research Focus} 
\noindent Project workscopes are summarized. 
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
    \item\textit{Title (Lead - )}
        \begin{itemize}[topsep=-1ex,itemsep=-1ex,partopsep=1ex,parsep=1ex]
            \item 
            \item 
        \end{itemize}
    \item\textit{}
        \begin{itemize}[topsep=-1ex,itemsep=-1ex,partopsep=1ex,parsep=1ex]
            \item 
            \item 
        \end{itemize}
    \item\textit{}
        \begin{itemize}[topsep=-1ex,itemsep=-1ex,partopsep=1ex,parsep=1ex]
            \item 
            \item  
        \end{itemize}
    \item\textit{Conclusions - Crosscutting issues (Full team)}
        \begin{itemize}[topsep=-1ex,itemsep=-1ex,partopsep=1ex,parsep=1ex]
            \item Identify lessons learned.
            \item Develop future research pathways.
        \end{itemize}
\end{enumerate}

\subsection*{Importance and Relevance to Objectives}  
\paragraph*{Background.} 
\subparagraph*{}

\subparagraph*{}

\subparagraph*{}

\paragraph*{}

\paragraph*{}

\subparagraph*{}

\subparagraph*{}

\paragraph*{Unique project features.}

\subsection*{Logical Path, Workscope, Description of Tasks} 
\paragraph*{Workscope 1 - Title.} 
\subparagraph*{Task I. Title.}

\subparagraph*{Task II. Title.}

\subparagraph*{Task III. Title.}

\subparagraph*{Task IV. Title.}

\paragraph*{Workscope 2 - Title.} 
\subparagraph*{Task V. Title.}

\paragraph*{Workscope 6. Conclusions.}
\subparagraph*{Task N. Crosscutting Issues.} 
\noindent Lessons learned will be identified for project outcomes. Selected additional discussion will cover, but not limited to, the following -
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
    \item 
    \item  
    \item 
\end{enumerate}

\paragraph*{Logical Path to Work Accomplishment. Major Deliverables and Outcomes.} 
\noindent The following outcomes are anticipated -
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
    \item
    \item
    \item 
    \item
    \item
\end{enumerate}

%\noindent\textbf{Estimated Cost of the Project.} The maximum amount of \$400,000 is requested for the 2 year time period at \$200,000 per year.
%

\begin{longtable}{|L{0.27\linewidth}|c|c|c|c|c|c|c|c|c|c|c|c|}
    \caption*{\textbf{Timeframe for Execution of Proposed Project. Schedule, Roles, and Responsibilities.}}
    \label{tab-timeframe}\\
        \hline
        \multicolumn{1}{|c|}{\multirow{2}{*}{\textbf{TASKS}}}& 
        \multicolumn{4}{|c|}{\textbf{Y1}}&
        \multicolumn{4}{|c|}{\textbf{Y2}}&
        \multicolumn{4}{|c|}{\textbf{Y3}}
        \\
        \cline{2-13}
        &
        \textbf{Q1}& 
        \textbf{Q2}& 
        \textbf{Q3}& 
        \textbf{Q4}& 
        \textbf{Q1}& 
        \textbf{Q2}& 
        \textbf{Q3}& 
        \textbf{Q4}& 
        \textbf{Q1}& 
        \textbf{Q2}& 
        \textbf{Q3}& 
        \textbf{Q4}
        \\
        \hline
        I. 
        &\x\textbf{I} 
        &
        &
        &
        &
        &
        &
        &
        &
        &
        &
        &
        \\
        \hline
        II. 
        &\x\textbf{II} 
        &\x\textbf{II} 
        &
        &
        &
        &
        &
        &
        &
        & 
        & 
        & 
        \\
        \hline
        III. 
        &\x\textbf{III} 
        &
        &
        &
        &
        &
        &
        &
        &
        &
        & 
        & 
        \\
        \hline
        IV. 
        & 
        &\x\textbf{IV} 
        & 
        &
        &
        &
        &
        &
        &
        &
        &
        & 
        \\
        \hline
        V. 
        & 
        &\x\textbf{V} 
        &\x\textbf{V} 
        &\x\textbf{V}
        &
        &
        &
        &
        &
        &
        &
        &
        \\
        \hline
        VI. 
        & 
        & 
        & 
        &\x\textbf{VI}
        & 
        &
        &
        &
        &
        & 
        &
        &
        \\
        \hline
        VII. 
        & 
        & 
        & 
        &\x\textbf{VII}
        & 
        &
        &
        &
        &
        & 
        &
        &
        \\
        \hline
        VIII. 
        & 
        & 
        & 
        & 
        &\x\textbf{VIII} 
        & 
        &
        &
        &
        &
        &
        &
        \\
        \hline
        IX. 
        & 
        & 
        & 
        &\x\textbf{IX} 
        &\x\textbf{IX} 
        & 
        &
        &
        &
        &
        &
        &
        \\
        \hline
        X. 
        & 
        & 
        &
        &
        &
        &
        & 
        & 
        &\x\textbf{X} 
        &\x\textbf{X} 
        &\x\textbf{X}
        &\x\textbf{X}
        \\
        \hline
        XI. Crosscutting issues 
        & 
        & 
        &
        &
        &
        &
        & 
        & 
        & 
        &
        &\x\textbf{VII}
        &\x\textbf{VII} 
        \\
        \hline
\end{longtable}

\vspace{-\baselineskip}

\begin{longtable}{|l|l|l|}
    \caption*{\textbf{Milestones.}}
    \label{tab-milestones}\\
        \hline
        \multicolumn{1}{|c|}{\textbf{Milestone}}
        &\multicolumn{1}{|c|}{\textbf{Deliverable}}
        &\multicolumn{1}{|c|}{\textbf{Completion}}\\
        \hline
        \begin{tabular}[c]{@{}l@{}}
        0. Project\\logistics
        \end{tabular}
        &
        \begin{tabular}[c]{@{}l@{}}
        0.1. Kick-off report, roles, and responsibilities.\\
        0.2. Annual reports.\\
        0.3. Agency administrative requirements.\\
        0.4. Final project report.\\
        0.5. Journal paper submission.\\
        0.6. Professional society conference presentations.
        \end{tabular}
        & 
        \begin{tabular}[c]{@{}l@{}}
        0.1. Y1-Q2\\
        0.2. Annually\\
        0.3. Quarterly\\
        0.4. Y2-Q4\\
        0.5. Annually\\
        0.6. Annually
        \end{tabular}
        \\
        \hline
        \begin{tabular}[c]{@{}l@{}}
        1. 
        \end{tabular}
        &
        \begin{tabular}[c]{@{}l@{}}
        1.1.  \\
        1.2. 
        \end{tabular}
        &
        \begin{tabular}[c]{@{}l@{}}
        1.1.  \\
        1.2.  
        \end{tabular}
        \\
        \hline
        \begin{tabular}[c]{@{}l@{}}
        2. 
        \end{tabular}
        &
        \begin{tabular}[c]{@{}l@{}}
        2.1. \\
        2.2. \\
        2.3. 
        \end{tabular}
        &
        \begin{tabular}[c]{@{}l@{}}
        2.1. \\
        2.2. \\
        2.3. 
        \end{tabular}
        \\
        \hline
        \begin{tabular}[c]{@{}l@{}}
        3. 
        \end{tabular}
        &
        \begin{tabular}[c]{@{}l@{}}
        3.1. 
        \end{tabular}
        &
        \begin{tabular}[c]{@{}l@{}}
        3.1. 
        \end{tabular}
        \\
        \hline
        \begin{tabular}[c]{@{}l@{}}
        4. 
        \end{tabular}
        &
        \begin{tabular}[c]{@{}l@{}}
        4.1. \\
        4.2. \\
        4.3. 
        \end{tabular}
        &
        \begin{tabular}[c]{@{}l@{}}
        4.1. \\
        4.2. 
        \end{tabular}
        \\
        \hline
        \begin{tabular}[c]{@{}l@{}}
        5. 
        \end{tabular}
        &
        \begin{tabular}[c]{@{}l@{}}
        5.1. \\
        5.2. \\
        5.3. \\
        5.4. \\
        5.5. 
        \end{tabular}
        & 
        \begin{tabular}[c]{@{}l@{}}
        5.1. \\
        5.2. \\
        5.3. \\
        5.4. \\
        5.5. 
        \end{tabular}
        \\
        \hline
        \begin{tabular}[c]{@{}l@{}}
        6. 
        \end{tabular}
        &
        \begin{tabular}[c]{@{}l@{}}
        6.1. \\
        6.2.  
        \end{tabular}
        &
        \begin{tabular}[c]{@{}l@{}}
        6.1. \\
        6.2. 
        \end{tabular}
        \\
        \hline
        \begin{tabular}[c]{@{}l@{}}
        7. 
        \end{tabular}
        &
        \begin{tabular}[c]{@{}l@{}}
        7.1. 
        \end{tabular}
        &
        \begin{tabular}[c]{@{}l@{}}
        7.1. 
        \end{tabular}
        \\
        \hline
        \begin{tabular}[c]{@{}l@{}}
        8. 
        \end{tabular}
        &
        \begin{tabular}[c]{@{}l@{}}
        8.1. \\
        8.2. \\
        8.3. 
        \end{tabular}
        &
        \begin{tabular}[c]{@{}l@{}}
        8.1. \\
        8.2. \\
        8.3. 
        \end{tabular}
        \\
        \hline
        \begin{tabular}[c]{@{}l@{}}
        9. 
        \end{tabular}
        &
        \begin{tabular}[c]{@{}l@{}}
        9.1. \\
        9.2. \\
        9.3. 
        \end{tabular}
        &
        \begin{tabular}[c]{@{}l@{}}
        9.1. \\
        9.2. \\
        9.3. 
        \end{tabular}
        \\
        \hline
        \begin{tabular}[c]{@{}l@{}}
        10. 
        \end{tabular}
        &
        \begin{tabular}[c]{@{}l@{}}
        10.1. \\
        10.2. \\
        10.3. \\
        10.4. \\
        10.5. \\
        10.6. 
        \end{tabular}
        &
        \begin{tabular}[c]{@{}l@{}}
        10.1. \\
        10.2. \\
        10.3. \\
        10.4. \\
        10.5. \\
        10.6. 
        \end{tabular}
        \\
        \hline
        \begin{tabular}[c]{@{}l@{}}
        11. Discuss\\crosscutting\\issues
        \end{tabular}
        &
        \begin{tabular}[c]{@{}l@{}}
        11.1. Identify lessons learned.\\
        11.2. Develop future research and educational pathways.
        \end{tabular}
        &
        \begin{tabular}[c]{@{}l@{}}
        11.1. Y3-Q3\\
        11.2. Y3-Q4
        \end{tabular}
        \\
        \hline
\end{longtable}

\paragraph*{Contribution to advancing state of the art in a, b, and c.}
\noindent This research plan and deliverables will advance the state-of-the-art in a, b, and, c as it is expected to - (1), (2), (3), and (4).

\paragraph*{Description of facilities to be utilized to execute scope.}

\paragraph*{Challenges to accomplish tasks and planned mitigations.} 
\noindent The PI is based at a United States Research University. The co-PIs are based at \textit{Affiliations}. This proposed project supports the goals and research interests of the institutions in the field of \textit{major fields}.

\vspace{-\baselineskip}

\begin{longtable}{|L{0.45\linewidth}|L{0.45\linewidth}|}
    \caption*{}
    \label{tab-risk}\\
        \hline
        \multicolumn{1}{|c|}{\textbf{Challenges to accomplish tasks}}
        &\multicolumn{1}{|c|}{\textbf{Planned mitigations}}\\
        \hline

        & 
        \\
        \hline

        &  
        \\
        \hline

        &  
        \\
        \hline

        & 
        \\
        \hline
\end{longtable}

\paragraph*{Information, data, and plans.} 
\noindent All information has been presented in the technical scope.

\paragraph*{Quality assurance.} 
\noindent Our project team have practiced a QA approach, consistent with the graded approach established by the DOE Technical Integration Office, per the NEUP website and any additional requirements deemed necessary during contracting. All projects undergo safety review by experts. We have followed this QA expectation in all ongoing awards.

\paragraph*{Mandatory requirements.} 
\noindent We abide by current and pending information on nuclear-related federal funding contained in Form 3204. The Mandatory, Go/No-go, requirements will be coordinated with the institutional Office of Sponsored Research, the general counsel, and the Office of Technology Transfer. Each PI noted below will assume responsibility as the point of contact.

\vspace{-\baselineskip}

\begin{longtable}{|c|c|c|c|}
    \caption*{}
    \label{tab-mandatory-requirements}\\
        \hline
        \textbf{No.}
        &\textbf{Requirement}
        &\textbf{Responsibility}
        &\textbf{Evaluation}\\
        \hline
        1
        &Commitment to reporting and budget requirements
        &All PIs
        &Go/No-go 
        \\
        \hline
        2
        &10 CFR 851 Worker Safety and Health Program
        &All PIs
        &Go/No-go 
        \\
        \hline
        3
        &Export Control
        &All PIs
        &Go/No-go 
        \\
        \hline
        4
        &Standard Research Subcontract
        &All PIs
        &Go/No-go 
        \\
        \hline
        5
        &Quality Assurance
        &All PIs
        &Go/No-go 
        \\
        \hline
        6
        &Commitment to prepare additional contract elements
        &All PIs
        &Go/No-go 
        \\
        \hline
\end{longtable}

\vspace{-0.50\baselineskip}

\noindent\textbf{References}
\bibliographystyle{neup}
\setlength{\bibhang}{0pt}
\bibliography{references}

\end{document}