### 2022-08-07 presenations/general/general.tex
latex presentation general

\documentclass[aspectratio=1610,pdftex,dvipsnames,compress]{beamer}
\usetheme{Boadilla}
\usecolortheme{seahorse}
\beamertemplatenavigationsymbolsempty

\addtobeamertemplate{footnote}{\hskip -2em}{} %pushes footnote to margin

\setbeamerfont{title}{series=\bfseries}

\setbeamertemplate{page number in head/foot}[framenumber] %Just gives slide number; comment out for 1/7, 2/7, etc.

\definecolor{BackGround}{RGB}{255,250,240}
\setbeamercolor{background canvas}{bg=BackGround}

\usepackage{caption}
\captionsetup[figure]{labelformat=empty}

%packages and definitions
\usepackage{enumitem}
\setlist[itemize]{label=\textbullet}
\usepackage{nicefrac}
\usepackage{ulem}
\usepackage{graphicx,animate}
\usepackage{mypythonhighlight}% allows inclusion of graphics
\graphicspath{{../img/}} %path to graphics
\usepackage[yyyymmdd]{datetime} %date format
\renewcommand{\dateseparator}{.}

%%%%%
\usepackage{pgf}
\usepackage{tikz} % required for drawing custom shapes
\usetikzlibrary{shapes,arrows,automata,trees}
%%%%%

\usepackage{booktabs} % nice rules (thick lines) for tables
\usepackage{microtype} % improves typography for PDF

\usepackage[acronym,nomain,nonumberlist]{glossaries}
\makenoidxglossaries

\usepackage{xcolor,colortbl} %change font color
\usepackage[numbers,sort&compress]{natbib} 
\usepackage{bibentry}

\setlength{\bibsep}{0pt} %sets space between references
%\renewcommand{\bibsection}{} %suppresses large 'references' heading
\renewcommand\bibpreamble{\vspace{-0.2\baselineskip}} %sets spacing after heading if not using default references heading

%\begin{filecontents}{references.bib}
%\end{filecontents}

%%%%% user commands
\newcommand\blfootnote[1]{%
  \begingroup
  \renewcommand\thefootnote{}\footnote{#1}%
  \addtocounter{footnote}{-1}%
  \endgroup
}

\makeatletter
\renewcommand{\@biblabel}[1]{#1.\hfill} %bibliography ordered list has numbers left flush
\makeatother

\newcommand{\edit}[1]{\textcolor{blue}{#1}} %shortcut for changing font color on revised text
\newcommand{\fn}[1]{\footnote{#1}} %shortcut for footnote tag
\newcommand*\sq{\mathbin{\vcenter{\hbox{\rule{.3ex}{.3ex}}}}} %makes a small square as a separator $\sq$
\newcommand{\sk}[1]{\sout{#1}} %shortcut for strikethrough
\newcommand{\x}{\cellcolor{lightgray}} %use to shade in table cell

\newcommand{\acf}{\acrfull} %full acronym
\newcommand{\acl}{\acrlong} %long acronym
\newcommand{\acs}{\acrshort} %short acronym
\newcommand{\acfp}{\acrfullpl} %full acronym plural
\newcommand{\aclp}{\acrlongpl} %long acronym plural
\newcommand{\acsp}{\acrshortpl} %short acronym plural
\newcommand{\Acf}{\Acrfull} %full acronym first letter capital
\newcommand{\Acl}{\Acrlong} %long acronym first letter capital
%%%%%

%%%%%
\newacronym{nrs}{NRHES}{Nuclear Renewable Hybrid Energy System}
\newacronym{ahp}{AHP}{Analytical Hierarchy Process}
\newacronym{inl}{INL}{Idaho National Laboratory}
\newacronym{orl}{ORNL}{Oak Ridge National Laboratory}
\newacronym{anl}{ANL}{Argonne National Laboratory}
\newacronym{npp}{NPP}{Nuclear Power Plant}
\newacronym{smr}{SMR}{Small Modular Reactor}
\newacronym{ump}{UAMPS}{Utah Associated Municipal Power Systems}
\newacronym{nus}{NuScale}{NuScale Power, LLC}
\newacronym{nrc}{NRC}{United States Nuclear Regulatory Commission}
\newacronym{epri}{EPRI}{Electric Power Research Institute}
\newacronym{nerc}{NERC}{North American Electric Reliability Corporation}
\newacronym{ci}{CI}{Consistency Index}
\newacronym{cr}{CR}{Consistency Ratio}
\newacronym{htse}{HTSE}{High Temperature Steam Electrolysis}
\newacronym{lwr}{LWR}{Light Water Reactor}
\newacronym{eia}{EIA}{U.S. Energy Information Administration}
\newacronym{oer}{OER}{Online Educational Resource}
\newacronym{lms}{LMS}{Learning Management System}
\newacronym{cps}{CPS}{Cyber-Physical Systems}
\newacronym{nsf}{NSF}{National Science Foundation}
\newacronym{wsc}{WSC}{Western Services Corporation}
\newacronym{cae}{CAES}{Center for Advanced Energy Studies}
\newacronym{hsl}{HSSL}{Human System Simulation Laboratory}
\newacronym{pwr}{PWR}{Pressurized Water Reactor}
\newacronym{roi}{ROI}{Return on Investment}
\newacronym{ic}{I\&C}{Instrumentation \& Controls}
\newacronym{mwe}{MWe}{Megawatts-electric}
\newacronym{ics}{ICS}{Industrial Control Systems}
\newacronym{sca}{SCADA}{Supervisory Control and Data Acquisition}
\newacronym{ip}{IP}{Internet Protocol}
\newacronym{udp}{UDP}{User Datagram Protocol}
\newacronym{tva}{TVA}{Tennessee Valley Authority}
\newacronym{plc}{PLC}{Programmable Logic Controller}
\newacronym{vfd}{VFD}{Variable Frequency Drive}
\newacronym{khp}{KHNP}{Korean Hydro \& Nuclear Power Co., Ltd}
\newacronym{onl}{ORNL}{Oak Ridge National Laboratory}
\newacronym{jcp}{JCPOA}{Joint Comprehensive Plan of Action}
\newacronym{mim}{MITM}{Man in the Middle}
\newacronym{dos}{DDoS}{Distributed Denial of Service}
\newacronym{ds}{DoS}{denial-of-service}
\newacronym{MM}{MiTM}{Man in the Middle}
\newacronym{tcp}{TCP/IP}{Transmission Control Protocol/Internet Protocol}
\newacronym{dnp}{DNP3}{Distributed Network Protocol 3}
\newacronym{pra}{PRA}{Probabilistic Risk Assessment}
\newacronym{cs}{CS}{Critical System}
\newacronym{loca}{LOCA}{Loss of Coolant Accident}
\newacronym{hmi}{HMI}{Human Machine Interface}
\newacronym{pha}{PHA}{Preliminary Hazards Analysis}
\newacronym{bol}{BOL}{beginning-of-life}
\newacronym{eol}{EOL}{end-of-life}
\newacronym{mol}{MOL}{middle-of-life}
\newacronym{imu}{IMUNES}{Integrated Multiprotocol Network Emulator/Simulator}
\newacronym{ccc}{CCC}{Computing Community Consortium}
\newacronym{neu}{NEUP}{Nuclear Energy University Program}
\newacronym{doe}{DOE}{United States Department of Energy}
\newacronym{nei}{NEI}{Nuclear Energy Institute}
\newacronym{nit}{NITRD}{Networking Information Technology Research \& Development Program}
\newacronym{rcs}{RCS}{Reactor Cooling System}
\newacronym{generic}{GPWR}{Generic Pressurized Water Reactor}
\newacronym{pcap}{PCAP}{packet capture file}
%%%%%

\title[Short title for footer]{Long title for title page}

\author[Short author for footer]{
    Name\\
    \vspace{0.25in}
    Another name\\
    \vspace{0.05in}
    \includegraphics{img/} 
    \includegraphics{img/}
}

\institute[Short institute for footer]{
    \vspace{0.05in}
    Institution\\
    Subtitle\\
    \vspace{0.10in}
    Subtitle\\
    \vspace{0.15in}
    \includegraphics[width=0.20\textwidth]{logo}
    }

\date{\today}

\begin{document}

%%%%% title page with no footer
{
    \setbeamertemplate{footline}{} 
    \begin{frame}
        \titlepage
    \end{frame}
}

\begin{frame}{Title}
    \begin{enumerate}[label=(\arabic*),leftmargin=*,itemsep=7pt]
        \item .
        \item .
        \item .
    \end{enumerate}
    \vspace{0.25in}
    \begin{itemize}[leftmargin=*,itemsep=7pt]
        \item .
        \item . 
    \end{itemize}
\end{frame}

\begin{frame}{Title}
    \begin{itemize}[leftmargin=*,itemsep=7pt]
        \item .  
        \item .
        \item .
    \end{itemize}
\end{frame}

%\begin{frame}{Title}
%    \begin{itemize}
%            \only<1-2>{
%            \item .
%            }
%
%            \only<2>{
%            \item .
%            \item . \citep{citation}
%            }
%    \end{itemize}
%
%    \only<2>{
%        \blfootnote{\tiny\cite{citation}\tiny\bibentry{citation}}
%    }
%\end{frame}

%\begin{frame}{Title}
%    \begin{itemize}
%        \item . \citep{ref}
%        \item . \citep{ref}
%    \end{itemize}
%    \blfootnote{\tiny\cite{ref}\tiny\bibentry{ref}}
%    \blfootnote{\tiny\cite{ref}\tiny\bibentry{ref}}
%\end{frame}

%\begin{frame}{Title}
%    \begin{columns}
%        \begin{column}{0.75\textwidth}
%            \begin{figure}
%                \centering
%                \includegraphics[width=0.95\textwidth]{figure}
%                \caption{}
%            \end{figure}
%        \end{column}
%
%        \begin{column}{0.25\textwidth}
%            \begin{itemize}
%                \item .
%                \item .
%                \item .
%            \end{itemize}
%        \end{column}
%    \end{columns}
%\end{frame}

%\begin{frame}{Acknowledgements}
%    \centering
%    Research was funded by .
%\end{frame}

%\begin{frame}{References}
%    \bibliographystyle{neup}
%    \footnotesize
%    \bibliography{references}
%\end{frame}

\end{document}