### 2023-05-11 cv-templates/cv-template-research-academic.tex

%@TheDoctorRAB
%resume
%
%%%%% general 
\documentclass[11pt,letterpaper]{article}
\usepackage[lmargin=1in,rmargin=1in,tmargin=1in,bmargin=1in]{geometry}
\usepackage[pagewise]{lineno} %line numbering
\usepackage{setspace}
\usepackage{ulem} %strikethrough - do not \sout{\cite{}}
\usepackage[pdftex,dvipsnames]{xcolor,colortbl} %change font color
\usepackage{graphicx}
\usepackage{filecontents}
\usepackage{tablefootnote}
\usepackage{footnotehyper}
\usepackage{float}
\usepackage{mypythonhighlight,verbatim}
%\usepackage{subfig}
\usepackage[yyyymmdd]{datetime} %date format
\renewcommand{\dateseparator}{.}
\graphicspath{{../img/}} %path to graphics
\setcounter{secnumdepth}{5} %set subsection to nth level
\usepackage{needspace}
\usepackage[stable,hang,flushmargin]{footmisc} %footnotes in section titles and no indent; standard.bst
\usepackage[inline]{enumitem}
\usepackage{boldline}
\usepackage{makecell}
\usepackage{booktabs}
\usepackage{amssymb}
\usepackage{gensymb}
\usepackage{amsmath,nicefrac}
\usepackage{physics}
\usepackage{lscape}
\usepackage{array}
\usepackage{chngcntr}
\usepackage{hyperref}
\hypersetup{colorlinks,linkcolor=black,citecolor=black,urlcolor=blue} 
\usepackage{sectsty}
\usepackage{textcomp}
\usepackage{lastpage}
\usepackage{xargs} %for \newcommandx
\usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes} %makes colored boxes for commenting
\usepackage{soul}
\usepackage{color}
\usepackage{marginnote}
\usepackage[figure,table]{totalcount}
\usepackage[capitalise]{cleveref}
\usepackage{nameref}
\usepackage{bold-extra}
\usepackage{parskip}
%%%%%

%%%%% fonts
\usepackage{times}
%arial - uncomment next two lines
%\usepackage{helvet}
%\renewcommand{\familydefault}{\sfdefault}
%%%%%

%%%%% references
\usepackage{bibentry} %lists references without \citep
\usepackage[round,semicolon]{natbib} %for (Borrelli 2021; Clooney 2019) - standard.bst 
%\usepackage[numbers,sort&compress]{natbib} %for [1-3] - nsf.bst, neup.bst
%\setlength{\bibsep}{7pt} %sets space between references
%\renewcommand{\bibsection}{} %suppresses large 'references' heading
%\renewcommand\bibpreamble{\vspace{\baselineskip}} %sets spacing after heading if not using default references heading
%%%%%

%%%%% heading formatting
\usepackage[toc,title]{appendix}
\usepackage[acronym,nomain,nonumberlist]{glossaries}
\makenoidxglossaries
\usepackage{titlesec,titletoc}
%\renewcommand{\thepart}{ARTICLE \Roman{part}} %puts the label into the command so \thelabel will carry through
%\renewcommand{\thesection}{\arabic{section}} %puts the label into the command so \thelabel will carry through
%\titleformat{\part}{\normalfont\Large\bfseries\filcenter}{\thepart}{}{}[]
%\titlespacing*\part{0pt}{0.95\baselineskip}{0.75\baselineskip}

\titleformat{\section}{\normalfont\Large\bfseries\scshape}{\thesection}{}{}[]
\titlespacing*\section{0pt}{0.95\baselineskip}{0.15\baselineskip}

\titleformat{\subsection}{\normalfont\large\bfseries\scshape}{\thesubsection}{}{}[]
\titlespacing*\subsection{0pt}{0.55\baselineskip}{0.10\baselineskip}

\titleformat{\subsubsection}[runin]{\normalfont\normalsize\bfseries\scshape}{\thesubsubsection}{}{}[]
\titlespacing*\subsubsection{0pt}{0.30\baselineskip}{0.05\baselineskip}

\titleformat{\paragraph}[runin]{\normalfont\normalsize\itshape}{\theparagraph}{-1em}{}[]
\titlespacing*\paragraph{0pt}{0.05\baselineskip}{0.05\baselineskip}

%\titleformat{\subparagraph}[runin]{\normalfont\normalsize\itshape}{\thesubparagraph}{-1em}{}[.]
%\titlespacing*\subparagraph{0pt}{0.40\baselineskip}{0.25\baselineskip}
%\titleformat{\paragraph}[hang]{\normalfont\normalsize\bfseries}{\theparagraph}{5pt}{}[]
%\titlespacing*\paragraph{0pt}{0.50\baselineskip}{0.25\baselineskip}
%\titleformat{\subparagraph}[runin]{\normalfont\normalsize\itshape}{\thesubparagraph}{-1em}{}[.]
%\titlespacing*\subparagraph{0pt}{0.40\baselineskip}{0.20\baselineskip}
%%%%%

%%%%% editing
\newcommand{\edit}[1]{\textcolor{blue}{#1}} %shortcut for changing font color on revised text
\newcommand{\fn}[1]{\footnote{#1}} %shortcut for footnote tag
\newcommand*\sq{\mathbin{\vcenter{\hbox{\rule{.3ex}{.3ex}}}}} %makes a small square as a separator $\sq$
%\newcommand{\sk}[1]{\sout{#1}} %shortcut for default strikethrough - do not sk through citep
\newcommand\sk{\bgroup\markoverwith{\textcolor{red}{\rule[0.5ex]{1pt}{1pt}}}\ULon} %strikethrough with red line; not in \section{}
%\st{} does strikethrough using soul package but does not like acronyms
%%%%%

%%%%% colors
%http://latexcolor.com/
%https://en.wikibooks.org/wiki/LaTeX/Colors#:~:text=black%2C%20blue%2C%20brown%2C%20cyan,be%20available%20on%20all%20systems.
\definecolor{aliceblue}{rgb}{0.94, 0.97, 1.0}
\definecolor{antiquewhite}{rgb}{0.98, 0.92, 0.84}
\definecolor{lightmauve}{rgb}{0.86, 0.82, 1.0}
\definecolor{brilliantlavender}{rgb}{0.96, 0.73, 1.0}
\definecolor{brandeisblue}{rgb}{0.0, 0.44, 1.0}
\definecolor{darkmidnightblue}{rgb}{0.0, 0.2, 0.4}
%%%%%

%%%%% acronyms
\newcommand{\acf}{\acrfull} %full acronym
\newcommand{\acl}{\acrlong} %long acronym
\newcommand{\acs}{\acrshort} %short acronym

\newcommand{\acfp}{\acrfullpl} %full acronym plural
\newcommand{\aclp}{\acrlongpl} %long acronym plural
\newcommand{\acsp}{\acrshortpl} %short acronym plural
%%%%%

%%%%% header and footer
\usepackage{fancyhdr} 
\pagestyle{fancy}
\fancyhf{} %page number on bottom right
\lhead{\scriptsize Name}
\chead{\scriptsize \textit{Vita}}
\rhead{\scriptsize \today} %date
\rfoot{\thepage} %page number
%%%%%

%%%%% citations
\begin{filecontents}{journal.bib}
\end{filecontents}
\begin{filecontents}{book.bib}
\end{filecontents}
\begin{filecontents}{conference.bib}
\end{filecontents}
\begin{filecontents}{report.bib}
\end{filecontents}
\begin{filecontents}{presentation.bib}
\end{filecontents}
%%%%%

%%%%% acronyms
% alphabetical ordering is automated
\newacronym{nrs}{NRHES}{Nuclear Renewable Hybrid Energy System}
\newacronym{ahp}{AHP}{Analytical Hierarchy Process}
\newacronym{inl}{INL}{Idaho National Laboratory}
\newacronym{orl}{ORNL}{Oak Ridge National Laboratory}
\newacronym{anl}{ANL}{Argonne National Laboratory}
\newacronym{npp}{NPP}{Nuclear Power Plant}
\newacronym{smr}{SMR}{Small Modular Reactor}
\newacronym{ump}{UAMPS}{Utah Associated Municipal Power Systems}
\newacronym{nus}{NuScale}{NuScale Power, LLC}
\newacronym{nrc}{NRC}{United States Nuclear Regulatory Commission}
\newacronym{epri}{EPRI}{Electric Power Research Institute}
\newacronym{nerc}{NERC}{North American Electric Reliability Corporation}
\newacronym{ci}{CI}{Consistency Index}
\newacronym{cr}{CR}{Consistency Ratio}
\newacronym{htse}{HTSE}{High Temperature Steam Electrolysis}
\newacronym{lwr}{LWR}{Light Water Reactor}
\newacronym{eia}{EIA}{U.S. Energy Information Administration}
\newacronym{oer}{OER}{Online Educational Resource}
\newacronym{lms}{LMS}{Learning Management System}
\newacronym{cps}{CPS}{Cyber-Physical Systems}
\newacronym{nsf}{NSF}{National Science Foundation}
\newacronym{wsc}{WSC}{Western Services Corporation}
\newacronym{cae}{CAES}{Center for Advanced Energy Studies}
\newacronym{hsl}{HSSL}{Human System Simulation Laboratory}
\newacronym{pwr}{PWR}{Pressurized Water Reactor}
\newacronym{bwr}{BWR}{Boiling Water Reactor}
\newacronym{roi}{ROI}{Return on Investment}
\newacronym{ic}{I\&C}{Instrumentation \& Controls}
\newacronym{mwe}{MWe}{Megawatts-electric}
\newacronym{ics}{ICS}{Industrial Control Systems}
\newacronym{sca}{SCADA}{Supervisory Control and Data Acquisition}
\newacronym{ip}{IP}{Internet Protocol}
\newacronym{udp}{UDP}{User Datagram Protocol}
\newacronym{tva}{TVA}{Tennessee Valley Authority}
\newacronym{plc}{PLC}{Programmable Logic Controller}
\newacronym{vfd}{VFD}{Variable Frequency Drive}
\newacronym{khp}{KHNP}{Korean Hydro \& Nuclear Power Co., Ltd}
\newacronym{onl}{ORNL}{Oak Ridge National Laboratory}
\newacronym{jcp}{JCPOA}{Joint Comprehensive Plan of Action}
\newacronym{mim}{MITM}{Man in the Middle}
\newacronym{dos}{DDoS}{Distributed Denial of Service}
\newacronym{tcp}{TCP/IP}{Transmission Control Protocol/Internet Protocol}
\newacronym{dnp}{DNP3}{Distributed Network Protocol 3}
\newacronym{pra}{PRA}{Probabilistic Risk Assessment}
\newacronym{cs}{CS}{Critical System}
\newacronym{loc}{LOCA}{Loss of Coolant Accident}
\newacronym{hmi}{HMI}{Human Machine Interface}
\newacronym{pha}{PHA}{Preliminary Hazards Analysis}
\newacronym{bol}{BOL}{Beginning-of-Life}
\newacronym{eol}{EOL}{End-of-Life}
\newacronym{mol}{MOL}{Middle-of-Life}
\newacronym{imu}{IMUNES}{Integrated Multiprotocol Network Emulator/Simulator}
\newacronym{ccc}{CCC}{Computing Community Consortium}
\newacronym{neu}{NEUP}{Nuclear Energy University Program}
\newacronym{doe}{DOE}{United States Department of Energy}
\newacronym{nei}{NEI}{Nuclear Energy Institute}
\newacronym{nit}{NITRD}{Networking Information Technology Research \& Development Program}
\newacronym{rcs}{RCS}{Reactor Cooling System}
\newacronym{con}{IC}{Initial Condition}
\newacronym{csi}{CSIS}{Center for Strategic \& International Studies}
\newacronym{pcap}{PCAP}{packet capture file}
\newacronym{dc}{DC}{Direct-Current}
\newacronym{ac}{AC}{Alternating-Current}
\newacronym{iff}{UIIF}{Idaho Falls Center for Higher Education}
\newacronym{snl}{SNL}{Sandia National Laboratory}
\newacronym{cie}{CIE}{Cyber-Informed Engineering}
\newacronym{cds}{CRDS}{Control Rod Drive System}
\newacronym{cdm}{CRDM}{Control Rod Drive Mechanism}
\newacronym{fma}{FMEA}{Failure Modes \& Effects Analysis}
\newacronym{rpn}{RPN}{Risk Priority Number}
\newacronym{scr}{SCR}{silicon controller rectifier}
\newacronym{hvc}{HVAC}{Heating, Ventilation \& Air Conditioning}
\newacronym{ttb}{TTB}{Time-to-Boil}
\newacronym{sis}{SIS}{Safety Instrumented System}
\newacronym{usb}{USB}{Universal Serial Bus}
\newacronym{ack}{ACK}{Acknowledge}
\newacronym{seq}{SEQ}{Sequence}
\newacronym{npr}{NPR}{National Public Radio}
\newacronym{cyd}{CYDEF}{Cyber Defence QCD Corporation}
\newacronym{ara}{ARA}{Amsterdam-Rotterdam-Antwerp}
%\newacronym{}{}{}
%%%%%

\begin{document}

\nobibliography{journal,book,conference,report,presentation} %bib files
\bibliographystyle{standard} %custom bst file

{\centering
        \textsc{ %block caps font
            \textbf{
                Name\\
                {\footnotesize Title} \\
                {\footnotesize Insitution} \\
                {\footnotesize Department} \\
                {\footnotesize Address}\\
                {\footnotesize email $\sq$ Twitter $\sq$ web }
            }
        }
   \par}

\noindent\makebox[\linewidth]{\rule{\textwidth}{0.5pt}} %horizontal line spanning margins

\section*{Education}
\subsection*{Institution 1}
\subsubsection*{Doctor of Philosophy $\sq$ Major}\hfill{\textit{Month Year}}
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex,label=(\arabic*)]
    \item[]\textit{Dissertation Title}
    \item[]Examination fields -- 
    \item[]Committee -- 
\end{enumerate}

\subsection*{Institution 2}
\subsubsection*{Master of Science $\sq$ Major}\hfill{\textit{Month Year}}
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex,label=(\arabic*)]
    \item[]\textit{Thesis Title}
    \item[]Examination fields -- 
    \item[]Committee -- 
\end{enumerate}

\subsection*{Institution 3}
\subsubsection*{Bacheor of Science $\sq$ Major}\hfill{\textit{Month Year}}
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex,label=(\arabic*)]
    \item[]\textit{Capstone}
    \item[]Advisor -- 
\end{enumerate}

\section*{Research Experience}
\subsection*{Institution 1 $\sq$ Department}
\subsubsection*{Title}\hfill{\textit{Time period}}
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex,label=(\arabic*)]
    \item[]\indent\indent Other repsonsibilites, titles, positions \hfill{\textit{Time period}}
    \item[]\indent\indent Other repsonsibilites, titles, positions \hfill{\textit{Time period}}
    \item[]\indent\indent Other repsonsibilites, titles, positions \hfill{\textit{Time period}}
\end{enumerate}
\subsubsection*{Other Title}\hfill{\textit{Time period}}\\
Description

\section*{Related Research Experience}
\subsection*{Institution 1 $\sq$ Department}
\subsubsection*{Title}\hfill{\textit{Time period}}
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex,label=(\arabic*)]
    \item[]\indent\indent Other repsonsibilites, titles, positions \hfill{\textit{Time period}}
    \item[]\indent\indent Other repsonsibilites, titles, positions \hfill{\textit{Time period}}
    \item[]\indent\indent Other repsonsibilites, titles, positions \hfill{\textit{Time period}}
\end{enumerate}
\subsubsection*{Other Title}\hfill{\textit{Time period}}\\
Description

\section*{Related Professional Experience}
\subsection*{Institution 1 $\sq$ Department}
\subsubsection*{Title}\hfill{\textit{Time period}}
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex,label=(\arabic*)]
    \item[]\indent\indent Other repsonsibilites, titles, positions \hfill{\textit{Time period}}
    \item[]\indent\indent Other repsonsibilites, titles, positions \hfill{\textit{Time period}}
    \item[]\indent\indent Other repsonsibilites, titles, positions \hfill{\textit{Time period}}
\end{enumerate}
\subsubsection*{Other Title}\hfill{\textit{Time period}}\\
Description

\section*{Awarded Projects} %chronologically
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=.15ex,partopsep=1ex,parsep=1ex,label=(\arabic*)]
    \item Big proposal 1
    \item Big proposal 2
    \item Big proposal 3
\end{enumerate}

\section*{Equipment Acquisitions}
Fancy equipment

\section*{Thesis Committees}
\subsection*{Major Professor}
\subsubsection*{Current}
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex,label=(\arabic*)]
    \item[] Student
    \item[] Student
    \item[] Student
\end{enumerate}

\subsubsection*{Graduates}
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex,label=(\arabic*)]
    \item[] Student
    \item[] Student
    \item[] Student
\end{enumerate}

\subsection*{Committee Member}
\subsubsection*{Graduates}
\paragraph*{University 1}
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex,label=(\arabic*)]
    \item[] Student
    \item[] Student
    \item[] Student
\end{enumerate}

\subsubsection*{Current}
\paragraph*{University 1}
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex,label=(\arabic*)]
    \item[] Student
    \item[] Student
    \item[] Student
\end{enumerate}

\section*{Refereed Journal Publications}
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=.15ex,partopsep=1ex,parsep=1ex,label=(\arabic*)]
    \item Big paper
%    \item\bibentry{}
\end{enumerate}

\section*{Books \& Book Contributions}
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=.15ex,partopsep=1ex,parsep=1ex,label=(\arabic*)]
    \item Book chapter
%    \item\bibentry{}
\end{enumerate}

\section*{Refereed Conference Proceedings}
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=.15ex,partopsep=1ex,parsep=1ex,label=(\arabic*)]
    \item Big paper
%    \item\bibentry{}
\end{enumerate}

\section*{Non-Refereed Technical Reports}
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=.15ex,partopsep=1ex,parsep=1ex,label=(\arabic*)]
    \item Big paper
%    \item\bibentry{}
\end{enumerate}

\section*{Conference Presentations}
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=.15ex,partopsep=1ex,parsep=1ex,label=(\arabic*)]
    \item Big paper
%    \item\bibentry{}
\end{enumerate}

\section*{Teaching Experience}
\subsection*{Institution 1 $\sq$ Department}
\subsubsection*{Title}
\paragraph*{Course}\hfill{\textit{semester}}\\
Description

\subsection*{Institution 1 $\sq$ Department}
\subsubsection*{Title}
\paragraph*{Course}\hfill{\textit{semester}}\\
Description

\section*{Related Teaching Experience}
\subsection*{Institution 1 $\sq$ Department}
\subsubsection*{Title}
\paragraph*{Course}\hfill{\textit{semester}}
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex,label=(\arabic*)]
    \item[] Duties
\end{enumerate}
Description

\section*{Professional Service}
\subsection*{Institution 1}
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex,label=(\arabic*)]
    \item[] service things
\end{enumerate}

\subsection*{Academic Search Committees}
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex,label=(\arabic*)]
    \item[] committee
\end{enumerate}

\subsection*{Peer Reviewer}
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex,label=(\arabic*)]
    \item[] Journals, etc.
\end{enumerate}

\subsection*{Conference Committees}
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex,label=(\arabic*)]
    \item[] committee
\end{enumerate}

\subsection*{Technical Session Organizer}
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex,label=(\arabic*)]
    \item[] Session 
\end{enumerate}

\subsection*{Technical Session Chair}
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex,label=(\arabic*)]
    \item[] Session 
\end{enumerate}

\subsection*{Panels}
\subsubsection*{Moderator}
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex,label=(\arabic*)]
    \item[]Panel 
\end{enumerate}

\subsubsection*{Panelist}
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex,label=(\arabic*)]
    \item[]Panel 
\end{enumerate}

\subsection*{Workshops}
\subsubsection*{Workshop}\hfill{\textit{Time}}\\
Description

\subsection*{Invited Talks}
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex,label=(\arabic*)]
    \item[]Big talk\hfill{\textit{Date}}
\end{enumerate}

\subsection*{Symposia Organizer}
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex,label=(\arabic*)]
    \item[]Symposium 1\hfill{\textit{Date}}
\end{enumerate}

\subsection*{Community Service}
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex,label=(\arabic*)]
    \item[]Service\hfill{\textit{Date}}
\end{enumerate}

\subsection*{Outreach \& Related Professional Activities}
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex,label=(\arabic*)]
    \item[]Activity\hfill{\textit{Date}}
\end{enumerate}

\section*{Professional Training}
\subsection*{Event}
\subsubsection*{Role}
\paragraph*{Title}\hfill{\textit{Time}}\\
Description

\section*{Awards \& Fellowships}
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex,label=(\arabic*)]
    \item[]Award\hfill{\textit{Date}}
\end{enumerate}

\section*{Professional Societies \& Positions}
\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex,label=(\arabic*)]
    \item[]Member\hfill{\textit{Date}}
\end{enumerate}

\end{document}