### 2023-10-10 acronyms/acronyms.tex
acronyms

%put this in preamble
\usepackage[acronym,nomain,nonumberlist]{glossaries}
\makenoidxglossaries
%%%%% acronyms
% alphabetical ordering is automated
\newacronym{nrs}{NRHES}{Nuclear Renewable Hybrid Energy System}
\newacronym{ahp}{AHP}{Analytical Hierarchy Process}
\newacronym{inl}{INL}{Idaho National Laboratory}
\newacronym{orl}{ORNL}{Oak Ridge National Laboratory}
\newacronym{anl}{ANL}{Argonne National Laboratory}
\newacronym{npp}{NPP}{Nuclear Power Plant}
\newacronym{smr}{SMR}{Small Modular Reactor}
\newacronym{ump}{UAMPS}{Utah Associated Municipal Power Systems}
\newacronym{nus}{NuScale}{NuScale Power, LLC}
\newacronym{nrc}{NRC}{United States Nuclear Regulatory Commission}
\newacronym{epri}{EPRI}{Electric Power Research Institute}
\newacronym{nerc}{NERC}{North American Electric Reliability Corporation}
\newacronym{ci}{CI}{Consistency Index}
\newacronym{cr}{CR}{Consistency Ratio}
\newacronym{htse}{HTSE}{High Temperature Steam Electrolysis}
\newacronym{lwr}{LWR}{Light Water Reactor}
\newacronym{eia}{EIA}{U.S. Energy Information Administration}
\newacronym{oer}{OER}{Online Educational Resource}
\newacronym{lms}{LMS}{Learning Management System}
\newacronym{cps}{CPS}{Cyber-Physical Systems}
\newacronym{nsf}{NSF}{National Science Foundation}
\newacronym{wsc}{WSC}{Western Services Corporation}
\newacronym{cae}{CAES}{Center for Advanced Energy Studies}
\newacronym{hsl}{HSSL}{Human System Simulation Laboratory}
\newacronym{pwr}{PWR}{Pressurized Water Reactor}
\newacronym{roi}{ROI}{Return on Investment}
\newacronym{ic}{I\&C}{Instrumentation \& Controls}
\newacronym{mwe}{MWe}{Megawatts-electric}
\newacronym{ics}{ICS}{Industrial Control System}
\newacronym{sca}{SCADA}{Supervisory Control and Data Acquisition}
\newacronym{ip}{IP}{Internet Protocol}
\newacronym{udp}{UDP}{User Datagram Protocol}
\newacronym{tva}{TVA}{Tennessee Valley Authority}
\newacronym{plc}{PLC}{Programmable Logic Controller}
\newacronym{vfd}{VFD}{Variable Frequency Drive}
\newacronym{khp}{KHNP}{Korean Hydro \& Nuclear Power Co., Ltd}
\newacronym{onl}{ORNL}{Oak Ridge National Laboratory}
\newacronym{jcp}{JCPOA}{Joint Comprehensive Plan of Action}
\newacronym{mim}{MITM}{Man in the Middle}
\newacronym{dos}{DDoS}{Distributed Denial of Service}
\newacronym{tcp}{TCP/IP}{Transmission Control Protocol/Internet Protocol}
\newacronym{dnp}{DNP3}{Distributed Network Protocol 3}
\newacronym{pra}{PRA}{Probabilistic Risk Assessment}
\newacronym{cs}{CS}{Critical System}
\newacronym{loc}{LOCA}{Loss of Coolant Accident}
\newacronym{hmi}{HMI}{Human Machine Interface}
\newacronym{pha}{PHA}{Preliminary Hazards Analysis}
\newacronym{bol}{BOL}{beginning-of-life}
\newacronym{eol}{EOL}{end-of-life}
\newacronym{mol}{MOL}{middle-of-life}
\newacronym{imu}{IMUNES}{Integrated Multiprotocol Network Emulator}
\newacronym{ccc}{CCC}{Computing Community Consortium}
\newacronym{neu}{NEUP}{Nuclear Energy University Program}
\newacronym{doe}{DOE}{United States Department of Energy}
\newacronym{nei}{NEI}{Nuclear Energy Institute}
\newacronym{nit}{NITRD}{Networking Information Technology Research \& Development Program}
\newacronym{rcs}{RCS}{Reactor Cooling System}
\newacronym{nic}{NIC}{Network Interface Card}
\newacronym{gui}{GUI}{Graphical User Interface}
\newacronym{pc}{PCAP}{Packet Capture File}
\newacronym{dc}{DC}{direct-current}
\newacronym{ac}{AC}{alternating-current}
\newacronym{csi}{CSIS}{Center for Strategic \& International Studies}
\newacronym{ste}{STEM}{Science, Technology, Engineering, and Mathematics}
\newacronym{pi}{PI}{Principal Investigator}
\newacronym{ans}{ANS}{American Nuclear Society}
\newacronym{sac}{SACNAS}{Society for the Advancement of Chicanos/Hispanics and Native Americans in Science}
\newacronym{ore}{ORED}{Office of Research and Economic Development}
\newacronym{rfd}{RFD}{Office of Research and Faculty Development}
\newacronym{nbe}{NSBE}{National Society of Black Engineers}
\newacronym{inp}{INPO}{Institute of Nuclear Power Operations}
\newacronym{nan}{NANTeL}{National Academy for Nuclear Training e-Learning}
\newacronym{cda}{CDA}{Critical Digital Assets}
\newacronym{nca}{NCA-CD}{National Center of Excellence in Cyber Defense}
\newacronym{ttx}{TTX}{Tabletop Exercise}
\newacronym{cie}{CIE}{Cyber-Informed Engineering}
\newacronym{aas}{AAS}{Adversary-As-A-Service}
\newacronym{usa}{US}{United States}
\newacronym{iff}{UIIF}{Idaho Falls Center for Higher Education}
\newacronym{fy}{FY}{Fiscal Year}
%\newacronym{}{}{}
%%%%%
%put this in the document somewhere
\printnoidxglossary
%%%%