### 2024-06-04 acronyms/acronyms.tex
p2x

\newacronym{nrs}{NRHES}{Nuclear Renewable Hybrid Energy System}
\newacronym{ahp}{AHP}{Analytical Hierarchy Process}
\newacronym{inl}{INL}{Idaho National Laboratory}
\newacronym{orl}{ORNL}{Oak Ridge National Laboratory}
\newacronym{anl}{ANL}{Argonne National Laboratory}
\newacronym{smr}{SMR}{Small Modular Reactor}
\newacronym{nus}{NuScale}{NuScale Power, LLC}
\newacronym{epri}{EPRI}{Electric Power Research Institute}
\newacronym{ci}{CI}{Consistency Index}
\newacronym{cr}{CR}{Consistency Ratio}
\newacronym{lwr}{LWR}{Light Water Reactor}
\newacronym{oer}{OER}{Online Educational Resource}
\newacronym{cps}{CPS}{Cyber-Physical Systems}
\newacronym{wsc}{WSC}{Western Services Corporation}
\newacronym{hsl}{HSSL}{Human System Simulation Laboratory}
\newacronym{roi}{ROI}{Return on Investment}
\newacronym{mwe}{MWe}{Megawatts-electric}
\newacronym{sca}{SCADA}{Supervisory Control and Data Acquisition}
\newacronym{udp}{UDP}{User Datagram Protocol}
\newacronym{plc}{PLC}{Programmable Logic Controller}
\newacronym{khp}{KHNP}{Korean Hydro \& Nuclear Power Co., Ltd}
\newacronym{jcp}{JCPOA}{Joint Comprehensive Plan of Action}
\newacronym{mim}{MITM}{Man in the Middle}
\newacronym{tcp}{TCP/IP}{Transmission Control Protocol/Internet Protocol}
\newacronym{pra}{PRA}{Probabilistic Risk Assessment}
\newacronym{cs}{CS}{Critical System}
\newacronym{loc}{LOCA}{Loss of Coolant Accident}
\newacronym{pha}{PHA}{Preliminary Hazards Analysis}
\newacronym{imu}{IMUNES}{Integrated Multiprotocol Network Emulator}
\newacronym{ccc}{CCC}{Computing Community Consortium}
\newacronym{neu}{NEUP}{Nuclear Energy University Program}
\newacronym{doe}{DOE}{United States Department of Energy}
\newacronym{nei}{NEI}{Nuclear Energy Institute}
\newacronym{nit}{NITRD}{Networking Information Technology Research \& Development Program}
\newacronym{rcs}{RCS}{Reactor Cooling System}
\newacronym{nic}{NIC}{Network Interface Card}
\newacronym{pc}{PCAP}{Packet Capture File}
\newacronym{csi}{CSIS}{Center for Strategic \& International Studies}
\newacronym{ste}{STEM}{Science, Technology, Engineering, and Mathematics}
\newacronym{pi}{PI}{Principal Investigator}
\newacronym{ans}{ANS}{American Nuclear Society}
\newacronym{sac}{SACNAS}{Society for the Advancement of Chicanos/Hispanics and Native Americans in Science}
\newacronym{ore}{ORED}{Office of Research and Economic Development}
\newacronym{rfd}{RFD}{Office of Research and Faculty Development}
\newacronym{nbe}{NSBE}{National Society of Black Engineers}
\newacronym{inp}{INPO}{Institute of Nuclear Power Operations}
\newacronym{nan}{NANTeL}{National Academy for Nuclear Training e-Learning}
\newacronym{cda}{CDA}{Critical Digital Assets}
\newacronym{nca}{NCA-CD}{National Center of Excellence in Cyber Defense}
\newacronym{ttx}{TTX}{Tabletop Exercise}
\newacronym{cie}{CIE}{Cyber-Informed Engineering}
\newacronym{aas}{AAS}{Adversary-As-A-Service}
\newacronym{iff}{UIIF}{Idaho Falls Center for Higher Education}
\newacronym{fy}{FY}{Fiscal Year}
\newacronym{eps}{EPSCoR}{Established Program to Stimulate Competitive Research}
\newacronym{nas}{NAS}{National Academies}
\newacronym{brc}{BRC}{Blue Ribbon Commission on America's Nuclear Future}
\newacronym{hlw}{HLW}{High-Level Radioactive Waste}
\newacronym{cbs}{CBS}{Consent-Based Siting}
\newacronym{wpp}{WIPP}{Waste Isolation Pilot Plant}
\newacronym{cfr}{CFR}{Code of Federal Regulations}
\newacronym{gao}{GAO}{Government Accountability Office}
\newacronym{npp}{NPP}{Nuclear Power Plant}
\newacronym{ump}{UAMPS}{Utah Associated Municipal Power Systems}
\newacronym{nrc}{NRC}{Nuclear Regulatory Commission}
\newacronym{nerc}{NERC}{North American Electric Reliability Corporation}
\newacronym{ferc}{FERC}{Federal Energy Regulatory Commission}
\newacronym{htse}{HTSE}{High Temperature Steam Electrolysis}
\newacronym{eia}{EIA}{U.S. Energy Information Administration}
\newacronym{lms}{LMS}{Learning Management System}
\newacronym{nsf}{NSF}{National Science Foundation}
\newacronym{cae}{CAES}{Center for Advanced Energy Studies}
\newacronym{pwr}{PWR}{Pressurized Water Reactor}
\newacronym{bwr}{BWR}{Boiling Water Reactor}
\newacronym{ic}{I\&C}{Instrumentation \& Controls}
\newacronym{ics}{ICS}{Industrial Control Systems}
\newacronym{ip}{IP}{Internet Protocol}
\newacronym{tva}{TVA}{Tennessee Valley Authority}
\newacronym{vfd}{VFD}{Variable Frequency Drive}
\newacronym{onl}{ORNL}{Oak Ridge National Laboratory}
\newacronym{dos}{DDoS}{Distributed Denial of Service}
\newacronym{dnp}{DNP3}{Distributed Network Protocol 3}
\newacronym{dra}{DRA}{Dynamic Risk Assessment}
\newacronym{hmi}{HMI}{Human Machine Interface}
\newacronym{bol}{BOL}{Beginning-of-Life}
\newacronym{eol}{EOL}{End-of-Life}
\newacronym{mol}{MOL}{Middle-of-Life}
\newacronym{con}{IC}{Initial Condition}
\newacronym{dc}{DC}{Direct-Current}
\newacronym{ac}{AC}{Alternating-Current}
\newacronym{snl}{SNL}{Sandia National Laboratory}
\newacronym{cds}{CRDS}{Control Rod Drive System}
\newacronym{cdm}{CRDM}{Control Rod Drive Mechanism}
\newacronym{fma}{FMEA}{Failure Modes \& Effects Analysis}
\newacronym{rpn}{RPN}{Risk Priority Number}
\newacronym{scr}{SCR}{silicon controller rectifier}
\newacronym{hvc}{HVAC}{Heating, Ventilation \& Air Conditioning}
\newacronym{ttb}{TTB}{Time-to-Boil}
\newacronym{sis}{SIS}{Safety Instrumented System}
\newacronym{cis}{CISF}{Consolidated Interim Storage Facility}
\newacronym{iea}{IEA}{International Energy Agency}
\newacronym{snf}{SNF}{Spent Nuclear Fuel}
\newacronym{cic}{CISAC}{Center for Interantional Security and Cooperation}
\newacronym{usa}{US}{United states}
\newacronym{nwa}{NWPA}{Nuclear Waste Policy Act}
\newacronym{one}{NE}{Office of Nuclear Energy}
\newacronym{cgc}{CGC}{Common Ground Consortium}
\newacronym{cop}{CoPL}{Communities of Place}
\newacronym{cor}{CoPR}{Communities of Practice}
\newacronym{coh}{CoH}{Communities of Historical Experience}
\newacronym{ej}{EJ}{Environmental Justice}
\newacronym{ite}{ITEK}{Indigenous Traditional Technical Ecological Knowledge}
\newacronym{iae}{IAEA}{International Atomic Energy Agency}
\newacronym{rfi}{RFI}{Request for Information}
\newacronym{sta}{STAND}{Siting Tool for Advanced Nuclear Development}
\newacronym{ors}{OR-SAGE}{Oak Ridge Siting Analysis for Power Generation Expansion}
\newacronym{gis}{GIS}{Geographic Information System}
\newacronym{mtu}{MTU}{Metric Ton of Uranium}
\newacronym{mth}{MTHM}{Metric Ton of Heavy Metal}
\newacronym{isa}{ISA}{Idaho Settlement Agreement}
\newacronym{tru}{TRU}{Transuranic Waste}
\newacronym{cej}{CEJEST}{Climate and Economic Justice Screening Tool}
\newacronym{eja}{EJAtlas}{Global Atlas of Environmental Justice}
\newacronym{ejs}{EJScreen}{Environmental Justice Screening and Mapping Tool}
\newacronym{fpz}{FPTZ}{Fastest Path to Zero}
\newacronym{tra}{WebTRAGIS}{Web-Based Transportation Routing Analysis Geographic Information System}
\newacronym{lit}{LITE}{Land-area Identification, Tagging, and Exploration}
\newacronym{str}{START}{Stakeholder Tool for Assessing Radioactive Transportation}
\newacronym{nri}{NRIC}{National Reactor Innovation Center}
\newacronym{sec}{SEC}{Site Evaluation Criteria}
\newacronym{ceq}{CEQ}{Council on Environmental Quality}
\newacronym{epa}{EPA}{Environmental Protection Agency}
\newacronym{unf}{UNF-ST\&DARDS}{Used Nuclear Fuel-Storage, Transportation \& Disposal Analysis Resource and Data System}
\newacronym{gui}{GUI}{Graphical User Interface}
\newacronym{fem}{FEMA}{Federal Emergency Management Agency}
\newacronym{rap}{RAPT}{Resilience Analysis and Planning Tool}
\newacronym{gid}{GEOID}{Geographic Identifier}
\newacronym{icr}{I-CREWS}{Idaho Community-Engaged Resilience for Energy-Water Systems}
\newacronym{ew}{E-W}{Energy-Water}
\newacronym{ml}{ML}{Machine Learning}
\newacronym{set}{SETS}{Socio-Environmental Technological System}
\newacronym{rii}{RII}{Research Infrastructure Improvement}
\newacronym{aoo}{AOO}{Anticipated Operational Occurrence}
\newacronym{dbe}{DBE}{Design Basis Event}
\newacronym{bbe}{BDBE}{Beyond Design Basis Event}
\newacronym{fta}{FTA}{Fault Tree Analysis}
\newacronym{eta}{ETA}{Event Tree Analysis}
\newacronym{haz}{HAZOP}{Hazard and Operability Analysis}
\newacronym{onr}{ONR}{Office of Naval Research}
\newacronym{ui}{UI}{University of Idaho}
\newacronym{isu}{ISU}{Idaho State University}
\newacronym{bsu}{BSU}{Boise State University}
\newacronym{skp}{SKP}{Senior Key Personnel}
\newacronym{bpc}{BPC}{Broadening Participation in Computing}
\newacronym{epi}{EPI}{Energy Policy Institute}
\newacronym{ym}{YM}{Yucca Mountain}
\newacronym{nwf}{NWF}{Nuclear Waste Fund}
\newacronym{aea}{AEA}{Atomic Energy Act}
\newacronym{aec}{AEC}{Atomic Energy Commission}
\newacronym{un}{UN}{United Nations}
\newacronym{des}{DES}{Discrete Event Simulation}
\newacronym{nem}{NEIM}{Nuclear Engineering \& Industrial Management}
\newacronym{erd}{ERDA}{Energy Research and Development Administration}
\newacronym{rd}{R\&D}{Research \& Development}
\newacronym{nts}{NTS}{Nevada Test Site}
\newacronym{nrm}{NRMSAC}{Nevada Radioactive Materials Storage Advisory Committee}
\newacronym{tmi}{TMI}{Three Mile Island}
\newacronym{nwt}{NWTS}{National Waste Terminal Storage}
\newacronym{irg}{IRG}{Interagency Review Group on Nuclear Waste Management}
\newacronym{kwh}{kWh}{Kilowatt-Hour}
\newacronym{eis}{EIS}{Environmental Impact Statement}
\newacronym{ocm}{OCRWM}{Office of Civilian Radioactive Waste Management}
\newacronym{dot}{DOT}{Department of Transportation}
\newacronym{mrs}{MRS}{Monitored Retrievable Storage}
\newacronym{nju}{NJUS}{Nome Joint Utility System}
\newacronym{irr}{IRR}{Internal Rate of Return}
\newacronym{tci}{TCI}{Total Capital Investment}
\newacronym{lce}{LCOE}{Levelized Cost of Electricity}
\newacronym{npv}{NPV}{Net Present Value}
\newacronym{lcs}{LCOS}{Levelized Cost of Storage}
\newacronym{lac}{LACE}{Levelized Avoided Cost of Electricity}
\newacronym{ece}{ECOE}{Effective Cost of Energy}
\newacronym{llp}{LOLP}{Loss of Load Probability}
\newacronym{ies}{IES}{Integrated Energy System}
\newacronym{nrl}{NREL}{National Renewable Energy Laboratory}
\newacronym{hop}{HOPP}{Hybrid Optimization and Performance Platform}
\newacronym{lch}{LCOH}{Levelized Cost of Hydrogen}
\newacronym{fom}{FOM}{Figure of Merit}
\newacronym{nep}{NEPA}{National Environmental Policy Act}
\newacronym{mca}{MC\&A}{Materials Control and Accountability}
\newacronym{rv}{R\&V}{Reproducibility \& Validation}
\newacronym{cce}{CCE}{Consequence-Driven Cyber-informed Engineering}
\newacronym{ic3}{IC3I}{Intermountain Center for the Cybersecurity of Critical Infrastructure}
\newacronym{p2x}{P2X}{Power-to-X}
%\newacronym{}{}{}
%Add \input{$ACRONYM/acronyms} in the preamble, where $ACRONYM is the /path/to/file/