### 2025-12-09 acronyms/acronyms.tex

%%%%% acronyms
%
% add \input{$ACRONYM/acronyms} in the preamble, where $ACRONYM is the /path/to/file/
% best practice is to pick a code of 2 or 3 letters
% loosely organized
% just to make it easier to read
%
% \newacronym{}{}{}
%
%%%%%
%
%%%%% universities, education
\newacronym{bsu}{BSU}{Boise State University}
\newacronym{byu}{BYU}{Brigham Young University}
\newacronym{cei}{CEI}{College of Eastern Idaho}
\newacronym{isu}{ISU}{Idaho State University} 
\newacronym{cos}{CSI}{College of Southern Idaho}
\newacronym{cwi}{CWI}{College of Western Idaho}
\newacronym{noc}{NIC}{North Idaho College}
\newacronym{ui}{UI}{University of Idaho}
\newacronym{um}{UM}{University of Michigan}
\newacronym{und}{UND}{University of North Dakota}
\newacronym{usu}{USU}{Utah State University}
\newacronym{utk}{UTK}{University of Tennessee--Knoxville}
\newacronym{wsu}{WSU}{Washington State University}
%
\newacronym{cae}{CAES}{Center for Advanced Energy Studies}
\newacronym{csn}{CSNR}{Center for Space Nuclear Research}
\newacronym{css}{CISSE}{Colloquium for Information Systems Security Education}
\newacronym{ece}{ECE}{Electrical \& Computer Engineering}
\newacronym{epi}{EPI}{Energy Policy Institute}
\newacronym{hrc}{HERC}{Higher Education Research Council}
\newacronym{ic3}{IC3I}{Intermountain Center for the Cybersecurity of Critical Infrastructure}
\newacronym{iff}{UIIF}{Idaho Falls Center for Higher Education}
\newacronym{igm}{IGEM}{Idaho Global Entrepreneurial Mission Initiative}
\newacronym{msm}{MSE}{Micron School of Materials Science and Engineering}
\newacronym{nem}{NEIM}{Nuclear Engineering \& Industrial Management}
\newacronym{ore}{ORED}{Office of Research and Economic Development}
\newacronym{rfd}{RFD}{Office of Research and Faculty Development}
\newacronym{sbe}{ISBE}{Idaho State Board of Education}
\newacronym{sde}{ISDE}{Idaho State Department of Education}
%
\newacronym{bs}{BS}{Bachelor of Science}
\newacronym{ms}{MS}{Master of Science}
\newacronym{phd}{Ph.D.}{Doctor of Philosophy}
\newacronym{gsi}{GSI}{Graduate Student Instructor}
\newacronym{gra}{GRA}{Graduate Research Assistant}
%%%%% national laboratories
\newacronym{anl}{ANL}{Argonne National Laboratory}
\newacronym{inl}{INL}{Idaho National Laboratory}
\newacronym{lnl}{LANL}{Los Alamos National Laboratory}
\newacronym{nrl}{NREL}{National Renewable Energy Laboratory}
\newacronym{onl}{ORNL}{Oak Ridge National Laboratory}
\newacronym{pnl}{PNNL}{Pacific Northwest National Laboratory}
\newacronym{snl}{SNL}{Sandia National Laboratory}
%
\newacronym{bea}{BEA}{Battelle Energy Alliance}
\newacronym{hsl}{HSSL}{Human System Simulation Laboratory}
\newacronym{nri}{NRIC}{National Reactor Innovation Center}
\newacronym{nrf}{NRF}{Naval Reactors Facility}
%%%%% government, societies, ngo 
\newacronym{ans}{ANS}{American Nuclear Society}
\newacronym{asc}{ASCE}{American Society of Civil Engineers}
\newacronym{brc}{BRC}{Blue Ribbon Commission on America's Nuclear Future}
\newacronym{ccc}{CCC}{Computing Community Consortium}
\newacronym{cdc}{CDC}{Center for Disease Control}
\newacronym{ceq}{CEQ}{Council on Environmental Quality}
\newacronym{cfr}{CFR}{Code of Federal Regulations}
\newacronym{cic}{CISAC}{Center for International Security and Cooperation}
\newacronym{csi}{CSIS}{Center for Strategic \& International Studies}
\newacronym{dhs}{DHS}{Department of Homeland Security}
\newacronym{dod}{DOD}{Department of Defense}
\newacronym{doe}{DOE}{United States Department of Energy}
\newacronym{dot}{DOT}{Department of Transportation}
\newacronym{eia}{EIA}{U.S. Energy Information Administration}
\newacronym{eo}{EO}{Executive Order}
\newacronym{epa}{EPA}{Environmental Protection Agency}
\newacronym{epri}{EPRI}{Electric Power Research Institute}
\newacronym{eps}{EPSCoR}{Established Program to Stimulate Competitive Research}
\newacronym{erd}{ERDA}{Energy Research and Development Administration}
\newacronym{ferc}{FERC}{Federal Energy Regulatory Commission}
\newacronym{fem}{FEMA}{Federal Emergency Management Agency}
\newacronym{g7}{G7}{Group of Seven}
\newacronym{gao}{GAO}{Government Accountability Office}
\newacronym{iae}{IAEA}{International Atomic Energy Agency}
\newacronym{iea}{IEA}{International Energy Agency}
\newacronym{iee}{IEEE}{Institute of Electrical and Electronics Engineers}
\newacronym{ihl}{IHLRWM}{International High Level Radioactive Waste Management}
\newacronym{inp}{INPO}{Institute of Nuclear Power Operations}
\newacronym{iso}{ISO}{International Organization for Standardization}
\newacronym{nan}{NANTeL}{National Academy for Nuclear Training e-Learning}
\newacronym{nas}{NAS}{National Academy of Sciences}
\newacronym{nbe}{NSBE}{National Society of Black Engineers}
\newacronym{nca}{NCA-CD}{National Center of Excellence in Cyber Defense}
\newacronym{nei}{NEI}{Nuclear Energy Institute}
\newacronym{nerc}{NERC}{North American Electric Reliability Corporation}
\newacronym{ngo}{NGO}{Non-Governmental Organization}
\newacronym{nit}{NITRD}{Networking Information Technology Research \& Development Program}
\newacronym{nis}{NIST}{National Institute of Standards and Technology}
\newacronym{nna}{NNSA}{National Nuclear Security Administration}
\newacronym{nrc}{NRC}{Nuclear Regulatory Commission}
\newacronym{nrm}{NRMSAC}{Nevada Radioactive Materials Storage Advisory Committee}
\newacronym{nss}{NASA}{National Aeronautics and Space Administration}
\newacronym{rdi}{REDI}{Regional Economic Development for Eastern Idaho}
\newacronym{sac}{SACNAS}{Society for the Advancement of Chicanos/Hispanics and Native Americans in Science}
\newacronym{usd}{USDA}{United States Department of Agriculture}
\newacronym{un}{UN}{United Nations}
\newacronym{usa}{US}{United States}
\newacronym{wpp}{WIPP}{Waste Isolation Pilot Plant}
%%%%% industry
\newacronym{khp}{KHNP}{Korean Hydro \& Nuclear Power Co., Ltd}
\newacronym{nau}{NARUC}{National Association of Regulatory Utility Commissioners}
\newacronym{nju}{NJUS}{Nome Joint Utility System}
\newacronym{nus}{NuScale}{NuScale Power, LLC}
\newacronym{tva}{TVA}{Tennessee Valley Authority}
\newacronym{ump}{UAMPS}{Utah Associated Municipal Power Systems}
\newacronym{wsc}{WSC}{Western Services Corporation}
%%%%% nuclear
\newacronym{ala}{ALARA}{As Low As Reasonably Achievable}
\newacronym{ap}{AP}{Activation Product}
\newacronym{atr}{ATR}{Advanced Test Reactor}
\newacronym{bol}{BOL}{Beginning-of-Life}
\newacronym{bwr}{BWR}{Boiling Water Reactor}
\newacronym{cbs}{CBS}{Collaborative-Based Siting}
%\newacronym{cbs}{CBS}{Consent-Based Siting}
\newacronym{cdm}{CRDM}{Control Rod Drive Mechanism}
\newacronym{cds}{CRDS}{Control Rod Drive System}
\newacronym{cdu}{CANDU}{Canada Deuterium Uranium Reactor}
%\newacronym{cis}{CISF}{Consolidated Interim Storage Facility}
\newacronym{cis}{CISF}{Consolidated Interim Storage Facility}
\newacronym{con}{IC}{Initial Condition}
\newacronym{cp1}{CP-1}{Chicago Pile-1}
\newacronym{dec}{DECON}{Decontamination \& Dismantlement}
\newacronym{dnd}{D\&D}{Decontamination \& Decommissioning}
\newacronym{eb1}{EBR-I}{Experimental Breeder Reactor I}
\newacronym{eb2}{EBR-II}{Experimental Breeder Reactor II}
\newacronym{ebs}{EBS}{Engineered Barrier System}
\newacronym{eol}{EOL}{End-of-Life}
\newacronym{fp}{FP}{Fission Product}
\newacronym{gtcc}{GTCC}{Greater than Class C}
\newacronym{hlw}{HLW}{High-Level Radioactive Waste}
\newacronym{hdap}{HDAP}{High-Dose Activation Product}
\newacronym{hop}{HOPP}{Hybrid Optimization and Performance Platform}
\newacronym{htse}{HTSE}{High Temperature Steam Electrolysis}
\newacronym{hvc}{HVAC}{Heating, Ventilation \& Air Conditioning}
\newacronym{ies}{IES}{Integrated Energy System}
\newacronym{isfsf}{ISFSF}{Interim Spent Fuel Storage Facility}
\newacronym{ke}{KE}{Kinetic Energy}
\newacronym{kmp}{KMP}{Key Measurement Point}
\newacronym{kwh}{kWh}{Kilowatt-Hour}
\newacronym{lft}{LFTR}{Liquid Fluoride Thorium Reactor}
\newacronym{llw}{LLW}{Low Level Waste}
\newacronym{loc}{LOCA}{Loss of Coolant Accident}
\newacronym{lwr}{LWR}{Light Water Reactor}
\newacronym{mca}{MC\&A}{Materials Control and Accountability}
\newacronym{mcp}{MCNP}{Monte Carlo N-Particle Transport Code}
\newacronym{mol}{MOL}{Middle-of-Life}
\newacronym{mox}{MOX}{Mixed Oxide}
\newacronym{mrm}{mrem}{millirem}
\newacronym{mse}{MSRE}{Molten Salt Reactor Experiment}
\newacronym{msr}{MSR}{Molten Salt Reactor}
\newacronym{mth}{MTHM}{Metric Ton of Heavy Metal}
\newacronym{mtu}{MTU}{Metric Ton of Uranium}
\newacronym{mwe}{MWe}{Megawatts-electric}
\newacronym{mrs}{MRS}{Monitored Retrievable Storage}
\newacronym{pwr}{PWR}{Pressurized Water Reactor}
\newacronym{npp}{NPP}{Nuclear Power Plant}
\newacronym{nrs}{NRHES}{Nuclear Renewable Hybrid Energy System}
\newacronym{nts}{NTS}{Nevada Test Site}
\newacronym{nwt}{NWTS}{National Waste Terminal Storage}
\newacronym{psd}{PSDAR}{Post Shutdown Decommissioning Activities Report}
\newacronym{pur}{PUREX}{Plutonium Uranium Reduction Extraction}
\newacronym{rcs}{RCS}{Reactor Cooling System}
\newacronym{rtg}{RTG}{Radioisotope Thermoelectric Generator}
\newacronym{saf}{SAFSTOR}{Safe Storage}
\newacronym{sec}{SEC}{Site Evaluation Criteria}
\newacronym{sfr}{SFR}{Sodium Fast Reactor}
\newacronym{sis}{SIS}{Safety Instrumented System}
\newacronym{snf}{SNF}{Spent Nuclear Fuel}
\newacronym{smr}{SMR}{Small Modular Reactor}
\newacronym{snm}{SNM}{Special Nuclear Material}
\newacronym{swu}{SWU}{Separative Work Unit}
\newacronym{tbp}{TBP}{Tributyl Phosphate}
\newacronym{tmi}{TMI}{Three Mile Island}
\newacronym{triga}{TRIGA}{Training, Research, Isotopes, General Atomics}
\newacronym{trt}{TREAT}{Transient Reactor Test Facility}
\newacronym{tru}{TRU}{Transuranic Waste}
\newacronym{ttb}{TTB}{Time-to-Boil}
\newacronym{wwg}{WWG}{Weight Window Generator}
\newacronym{vov}{VOV}{Variance of Variance}
\newacronym{ym}{YM}{Yucca Mountain}
%%%%% cybersecurity
\newacronym{aas}{AAS}{Adversary-As-A-Service}
\newacronym{ac}{AC}{Alternating-Current}
\newacronym{ai}{AI}{Artificial Intelligence}
\newacronym{cce}{CCE}{Consequence-Driven Cyber-informed Engineering}
\newacronym{cda}{CDA}{Critical Digital Assets}
\newacronym{cdo}{CD-0}{Critical Decision-0}
\newacronym{cie}{CIE}{Cyber-Informed Engineering}
\newacronym{cs}{CS}{Critical System}
\newacronym{cps}{CPS}{Cyber-Physical Systems}
\newacronym{dc}{DC}{Direct-Current}
\newacronym{ddml}{DDML}{Data Driven Machine Learning}
\newacronym{dnp}{DNP3}{Distributed Network Protocol 3}
\newacronym{dos}{DDoS}{Distributed Denial of Service}
\newacronym{dt}{DigitTwin}{Digital Twin}
\newacronym{ecs}{ECCS}{Emergency Core Cooling System}
\newacronym{gps}{GPS}{Global Positioning System}
\newacronym{gui}{GUI}{Graphical User Interface}
\newacronym{hitl}{HITL}{Human-in-the-Loop}
\newacronym{hmi}{HMI}{Human Machine Interface}
\newacronym{hpc}{HPC}{High Performance Computing}
\newacronym{hri}{HRI}{Human Risk Index}
\newacronym{ic}{I\&C}{Instrumentation \& Controls}
\newacronym{ics}{ICS}{Industrial Control Systems}
\newacronym{imu}{IMUNES}{Integrated Multiprotocol Network Emulator}
\newacronym{ip}{IP}{Internet Protocol}
\newacronym{mars}{MARS}{Maintenance of Advanced Reactor Sensors and Components}
\newacronym{mim}{MITM}{Man in the Middle}
\newacronym{ml}{ML}{Machine Learning}
\newacronym{nic}{NIC}{Network Interface Card}
\newacronym{p2x}{P2X}{Power-to-X}
\newacronym{pcp}{PCAP}{Packet Capture File}
\newacronym{plc}{PLC}{Programmable Logic Controller}
\newacronym{pp}{PP}{Physical Protection}
\newacronym{pr}{PR}{Proliferation Resistance}
\newacronym{rad}{RADICL}{Reconfigurable Attack-Defend Instructional Computing Laboratory}
\newacronym{rl}{RL}{Reinforcement Learning}
\newacronym{sca}{SCADA}{Supervisory Control and Data Acquisition}
\newacronym{scr}{SCR}{silicon controller rectifier}
\newacronym{tcp}{TCP/IP}{Transmission Control Protocol/Internet Protocol}
\newacronym{ttx}{TTX}{Tabletop Exercise}
\newacronym{udp}{UDP}{User Datagram Protocol}
\newacronym{vfd}{VFD}{Variable Frequency Drive}
\newacronym{xai}{XAI}{Explainable Artificial Intelligence}
%%%%% risk
\newacronym{ahp}{AHP}{Analytical Hierarchy Process}
\newacronym{ann}{ANN}{Artificial Neural Network}
\newacronym{aoo}{AOO}{Anticipated Operational Occurrence}
\newacronym{bbe}{BDBE}{Beyond Design Basis Event}
\newacronym{bn}{BN}{Bayesian Network}
\newacronym{cd}{CD}{Certificate of Deposit}
\newacronym{ci}{CI}{Consistency Index}
\newacronym{cr}{CR}{Consistency Ratio}
\newacronym{cvr}{CVaR}{Conditional Value at Risk}
\newacronym{des}{DES}{Discrete Event Simulation}
\newacronym{dra}{DRA}{Dynamic Risk Assessment}
\newacronym{dbe}{DBE}{Design Basis Event}
\newacronym{eco}{ECOE}{Effective Cost of Energy}
\newacronym{eta}{ETA}{Event Tree Analysis}
\newacronym{fma}{FMEA}{Failure Modes \& Effects Analysis}
\newacronym{fom}{FOM}{Figure of Merit}
\newacronym{fta}{FTA}{Fault Tree Analysis}
\newacronym{haz}{HAZOP}{Hazard and Operability Analysis}
\newacronym{hra}{HRA}{Human Reliability Analysis}
\newacronym{lac}{LACE}{Levelized Avoided Cost of Electricity}
\newacronym{lce}{LCOE}{Levelized Cost of Electricity}
\newacronym{lch}{LCOH}{Levelized Cost of Hydrogen}
\newacronym{lcs}{LCOS}{Levelized Cost of Storage}
\newacronym{llp}{LOLP}{Loss of Load Probability}
\newacronym{irr}{IRR}{Internal Rate of Return}
\newacronym{mtb}{MTBF}{Mean Time Before Failure}
\newacronym{mtf}{MTTF}{Mean Time To Failure}
\newacronym{npv}{NPV}{Net Present Value}
\newacronym{nrx}{NRI}{National Risk Index}
\newacronym{pdf}{PDF}{Probability Density Function}
\newacronym{pha}{PHA}{Preliminary Hazards Analysis}
\newacronym{pid}{P\&ID}{Piping \& Instrumentation Diagram}
\newacronym{pn}{PN}{Petri Net}
\newacronym{pra}{PRA}{Probabilistic Risk Assessment}
\newacronym{prm}{PRM}{Project Risk Management}
\newacronym{roi}{ROI}{Return on Investment}
\newacronym{rpn}{RPN}{Risk Priority Number}
\newacronym{sea}{SECA}{Simultaneous Evaluation of Criteria and Alternatives}
\newacronym{sra}{SRA}{Safety Risk Analysis}
\newacronym{tci}{TCI}{Total Capital Investment}
\newacronym{top}{TOPSIS}{Technique for Order of Preference by Similarity to Ideal Solution}
\newacronym{trp}{THERP}{Technique for Human Error Rate Prediction}
%%%%% education
\newacronym{ihe}{IHE}{Institution of Higher Education}
\newacronym{lms}{LMS}{Learning Management System}
\newacronym{oer}{OER}{Online Educational Resource}
\newacronym{ste}{STEM}{Science, Technology, Engineering, and Mathematics}
%%%%% proposals
\newacronym{ate}{ATE}{Advanced Technological Education}
\newacronym{ay}{AY}{Academic Year}
\newacronym{bpc}{BPC}{Broadening Participation in Computing}
\newacronym{cgc}{CGC}{Common Ground Consortium}
\newacronym{dcl}{DCL}{Dear Colleague Letter}
\newacronym{dmp}{DMP}{Data Management Plan}
\newacronym{fna}{F\&A}{Facilities and Administrative}
\newacronym{fte}{FTE}{Full Time Equivalent}
\newacronym{fy}{FY}{Fiscal Year}
\newacronym{icr}{I-CREWS}{Idaho Community-Engaged Resilience for Energy-Water Systems}
\newacronym{irb}{IRB}{Institutional Review Board}
\newacronym{its}{ITEST}{Innovative Technology Experiences for Students and Teachers}
\newacronym{mdc}{MTDC}{Modified Total Direct Cost}
\newacronym{mls}{MILES}{Managing Idaho's Landscapes for Ecosystem Services}
\newacronym{neu}{NEUP}{Nuclear Energy University Program}
\newacronym{nkn}{NKN}{Northwest Knowledge Network}
\newacronym{nsf}{NSF}{National Science Foundation}
\newacronym{one}{NE}{Office of Nuclear Energy}
\newacronym{onr}{ONR}{Office of Naval Research}
\newacronym{pi}{PI}{Principal Investigator}
\newacronym{rd}{R\&D}{Research \& Development}
\newacronym{rfi}{RFI}{Request for Information}
\newacronym{rii}{RII}{Research Infrastructure Improvement}
\newacronym{rv}{R\&V}{Reproducibility \& Validation}
\newacronym{skp}{SKP}{Senior Key Personnel}
\newacronym{trl}{TRL}{Technology Readiness Level}
\newacronym{vo}{CPS-VO}{CPS Virtual Organization}
%%%%% social science, policy
\newacronym{aea}{AEA}{Atomic Energy Act}
\newacronym{aec}{AEC}{Atomic Energy Commission}
\newacronym{afs}{AF}{Alternative Future}
\newacronym{aul}{AULG}{Affected Unit of Local Government}
\newacronym{caa}{CAA}{Consolidated Appropriations Act}
\newacronym{can}{CANT}{Citizens Against Nuclear Trash}
\newacronym{cej}{CEJEST}{Climate and Economic Justice Screening Tool}
\newacronym{coh}{CoH}{Communities of Historical Experience}
\newacronym{cop}{CoPL}{Communities of Place}
\newacronym{cor}{CoPR}{Communities of Practice}
\newacronym{eis}{EIS}{Environmental Impact Statement}
\newacronym{ej}{EJ}{Environmental Justice}
\newacronym{eja}{EJAtlas}{Global Atlas of Environmental Justice}
\newacronym{ejs}{EJScreen}{Environmental Justice Screening and Mapping Tool}
\newacronym{eoi}{EoI}{Expression of Interest}
\newacronym{ept}{EPAct}{Energy Policy Act}
\newacronym{ew}{E-W}{Energy-Water}
\newacronym{fpz}{FPTZ}{Fastest Path to Zero}
\newacronym{ges}{GESMO}{Generic Environmental Impact Statement for Mixed Oxide Fuel}
\newacronym{gid}{GEOID}{Geographic Identifier}
\newacronym{gis}{GIS}{Geographic Information System}
\newacronym{gsa}{GSA}{General Services Administration}
\newacronym{ida}{IDAPA}{Idaho Administrative Procedures Act}
\newacronym{ipc}{IPCC}{Intergovernmental Panel on Climate Change}
\newacronym{irg}{IRG}{Interagency Review Group on Nuclear Waste Management}
\newacronym{isa}{ISA}{Idaho Settlement Agreement}
\newacronym{ite}{ITEK}{Indigenous Traditional Technical Ecological Knowledge}
\newacronym{jcp}{JCPOA}{Joint Comprehensive Plan of Action}
\newacronym{lea}{LEA}{Local Educational Agency}
\newacronym{lit}{LITE}{Land-area Identification, Tagging, and Exploration}
\newacronym{nce}{NCES}{National Center for Education Statistics}
\newacronym{nep}{NEPA}{National Environmental Policy Act}
\newacronym{nwa}{NWPA}{Nuclear Waste Policy Act}
\newacronym{nwb}{NWTRB}{Nuclear Waste Technical Review Board}
\newacronym{nwf}{NWF}{Nuclear Waste Fund}
\newacronym{nwn}{NWN}{Office of the Nuclear Waste Negotiator}
\newacronym{nwr}{NWRPO}{Nuclear Waste Repository Project Office}
\newacronym{ocm}{OCRWM}{Office of Civilian Radioactive Waste Management}
\newacronym{omb}{OMB}{Office of Management and Budget}
\newacronym{ors}{OR-SAGE}{Oak Ridge Siting Analysis for Power Generation Expansion}
\newacronym{osa}{OSHA}{Occupational Safety and Health Administration}
\newacronym{osc}{OSCM}{Office of Standard Contract Management}
\newacronym{rap}{RAPT}{Resilience Analysis and Planning Tool}
\newacronym{set}{SETS}{Socio-Environmental Technological Systems}
\newacronym{sfs}{SFS}{Sustainable Future Scenarios}
\newacronym{sta}{STAND}{Siting Tool for Advanced Nuclear Development}
\newacronym{tpa}{TSPA}{Total System Performance Assessment}
\newacronym{tra}{WebTRAGIS}{Web-Based Transportation Routing Analysis Geographic Information System}
\newacronym{str}{START}{Stakeholder Tool for Assessing Radioactive Transportation}
\newacronym{unf}{UNF-ST\&DARDS}{Used Nuclear Fuel-Storage, Transportation \& Disposal Analysis Resource and Data System}
\newacronym{wef}{WEF}{World Economic Forum}
%%%%% water
\newacronym{adi}{ADI}{Agricultural Drought Index}
\newacronym{app}{AP}{Average Annual Rainfall Volume}
\newacronym{bws}{BWS}{Baseline Water Stress}
\newacronym{cq}{CI}{Consequence Subindex}
\newacronym{ds}{DS}{Drought Severity}
\newacronym{dta}{DTA}{Demand to Available}
\newacronym{edr}{EDR}{External Dependency Ratio}
\newacronym{efr}{EFR}{Environmental Flow Requirements}
\newacronym{ewr}{EWR}{Environmental Water Requirements}
\newacronym{exi}{ExI}{Exposure Indicator}
\newacronym{fri}{FRI}{Flood Risk Index}
\newacronym{gr}{GR}{Governance \& Regulation}
\newacronym{hi}{HI}{Drought Hazard Subindex}
\newacronym{hdi}{HDI}{Hydrological Drought Index}
\newacronym{kde}{KDE}{Kernel Density Estimation}
\newacronym{mdi}{MDI}{Meteorological Drought Index}
\newacronym{ndi}{NDI}{Normalized Deficit Index}
\newacronym{ndc}{NDC}{Normalized Deficit Cumulated}
\newacronym{rfr}{RFR}{Return Flow Ratio}
\newacronym{rwr}{RWR}{Renewable Water Resource}
\newacronym{sic}{SIC}{Stress Index Cumulated}
\newacronym{sii}{SII}{Stress Index Intra-annual}
\newacronym{ssd}{SS-DTA}{Sectoral and Statistical Demand to Availability}
\newacronym{sv}{SV}{Seasonal Variability}
\newacronym{vi}{VI}{Vulnerability Indicator}
\newacronym{wrc}{WRCC}{Water Resources Carrying Capacity}
\newacronym{wri}{WSRI}{Water Shortage Risk Index}
\newacronym{wsi}{W-ScaRI}{Water Scarcity Risk Index}
\newacronym{wta}{WTA}{Withdrawl To Availability}
%%%%% energy
\newacronym{iri}{IESRI}{International Energy Security Risk Index}
%%%%%
%\newacronym{}{}{}
%%%%%